### 2015-03-02 doc/report/survey/article.tex
survey report

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

%%\documentclass[]{spie}  %>>> use for US letter paper
\documentclass[12pt]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}
\usepackage[sort, numbers]{natbib}
\title{GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers} 
\author{Anirudh Jayakumar and Harshit Dokania \\
University of Illinois at Urbana-Champaign \\
\{ajayaku2,hdokani2\}@illinois.edu 
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

  \begin{document} 
  \maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
Today, many of the leading cloud service providers have geographically distributed data centers to serve millions of users around the world. The need for a geo-distributed data center arising from the need to provide low-latency and  high-availability of the different services. The workloads that run on these data centers are also highly varied. A typical data centers run jobs of different kinds including business-critical workloads---web servers, mail servers, instant messaging services etc, big-data processing, real-time data analytics and HPC jobs. As cloud services grow to serve more customers, there is an increasing need for a workload provisioning mechanism which can minimize the cost of operation of the data center while still obeying the SLA and also minimizing the user-perceived latency.

We present GeoSched, a job provisioning framework that is workload aware, energy-aware and cooling-aware. GeoSched considers the workload type (batch process, real-time processing, web-service), the energy source of the data center (brown and green energy), electricity pricing of the region, cooling techniques used in the data center (air economizer) and also the history of jobs to predict the load patterns and cluster utilization. We wish to evaluate GeoSched on industry workload traces and also on synthetic workloads that can mimic the real-world data center workload. 
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{sec:intro}  % \label{} allows reference to this section
Cost minimization has been extensively studied both in the context of a single data center and a geo-distributed data center. Most of the work has been targeted towards minimizing the energy consumption of the data center to reduce the operational cost. For example, the work by Gu et al. \cite{gu2014cost} studies the influence of task assignment, data placement  and data movement on the operational expenditure of large-scale geo-distributed data centers for big data applications. The authors jointly optimize  these three factors by proposing a 2-D Markov chain to derive the average task completion time and solve the model as a MILP problem. On similar lines Agarwal et al. \cite{agarwal2010volley} propose an automated data placement mechanism Volley for geo-distributed cloud services with the consideration of WAN bandwidth cost, data center capacity limits, data inter-dependencies, etc.

Yu et all.\cite{yu2015energy} propose minimizing energy cost for distributed Internet data centers (IDCs) in smart microgrids by considering system dynamics like uncertainties in electricity price, workload, renewable energy generation, and power outage state. They model the problem as a stochastic program and solve it using Lyapunov optimization technique. The work by Ren and He  \cite{ren2013coca} propose an online algorithm, called COCA to minimize data center operational cost while satisfying carbon neutrality. Unlike some of the existing research, COCA enables distributed server-level resource management: each server autonomously adjusts its processing speed and optimally decides the amount of workloads to process. This work is restricted to a single data center with heterogeneous resources.

Cooling energy is a substantial part of the total energy spent by a data center. According to some reports, this can be as high as 50\% \cite{sullivan2002alternating}, \cite{patel2003smart},\cite{sawyer2004calculating} of the total energy budget. There has been a lot of work both at the application/middleware level \cite{TempLDBSC11}, \cite{leverich2010energy} as well as provisioning techniques  \cite{tang2007thermal}, \cite{chen2010integrated} that try to minimize the cooling energy. Li et all. \cite{li2014coordinating} proposes SmartCool, a power optimization scheme that effectively coordinates different cooling techniques and dynamically manages workload allocation for jointly optimized cooling and server power. Unlike existing work that addresses different cooling techniques in an isolated manner, SmartCool integrates different cooling systems as a constrained optimization problem. Also, since geo-distributed data centers have different ambient temperatures, SmartCool dynamically dispatches the incoming requests among a network of data centers with heterogeneous cooling systems to best leverage the high efficiency of free cooling.

Other research related to cooling efficiency include the work by Liu et all \cite{liu2012renewable}  that reduces electricity cost and environmental impact using a holistic approach that integrates renewable supply, dynamic pricing, and cooling supply including chiller and outside air cooling, with IT workload planning to improve the overall sustainability of data center operations. The authors first predict renewable energy as well as IT demand. Then they use these predictions to generate an IT workload management plan that schedules IT workload and allocates IT resources within a data center according to time varying power supply and cooling efficiency. Pakbaznia and Pedram \cite{pakbaznia2009minimizing} focuses on minimizing power for both IT equipment and air conditioning power usage. The authors consider both server consolidation and task assignment together and formulate the resulting optimization problem a a ILP problem and present a heuristic algorithm that loves it in polynomial time.

There are also research done to look at the impact of failures in data centers on the operational cost. The work by Cui et al. \cite{cui2014shadows} considers the increase in energy consumption due to failures in large scale clouds. A slower shadow replication of the main process is used to recover from faults thereby saving energy upto 30\%. Guenter et al. \cite{guenter2011managing} also considers different trade-offs between cost, performance and reliability to formulate an optimization problem that minimizes energy costs, reliability costs and unmet demands. VM migration techniques are also used extensively to increase the energy efficiency of cloud data centers. One such approach is proposed by Rodero et al. \cite{rodero2012energy} where the author  propose to use applications۪s profiling (CPU, memory, bandwidth) to  design application-centric energy-aware strategy for VM allocation during VM migration.

Researchers have also studied the efficient use of green energy in data centers. Goiri et al. \cite{goiri2011greenslot} propose a system GreenSlot that maximizes the use of green energy available at data centers while ensuring all job deadlines are met. GreenSlot predicts the solar energy available in data centers in near future and schedules the workload to maximize green energy consumption and meeting jobs۪s deadlines. Green- Slot can increase green energy consumption by up to 117\% and decrease energy cost by up to 39\%, compared to a conventional scheduler.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem Statement}

Minimize the total cost of operations of a geo-distributed data center running cloud workloads by minimizing the energy costs subject to business constraints like the service level agreement(SLA) and other technical constraints like data center capacity limits. The energy costs will depend on the power usage of the IT infrastructure and also the energy required by the cooling infrastructure to maintain the desired temperature inside the data center. Additionally, the solution should also consider the different energy source(green and brown) available at the data center and the different workload types that are typical to a cloud data center.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Approaches }

We wish to consider all the major factors that impact the energy costs at the data center. Some of the factors and its impact are discussed in the below sub-section.

\subsection{Electricity pricing} 
We use the day-ahead hourly electricity prices to get the electricity cost of the data center which will enable us to provision workloads efficiently. Moreover, the prices in the day-ahead market fluctuate smoothly and can be predicted using time series forecasting techniques.
\subsection{Energy source in data centers}
To meet the energy requirements of data centers there is increasing adoption of renewable energy supply techniques (solar, wind). This has resulted in data centers being powered at least partially with green energy. We wish to effectively use the green energy to reduce costs. 
\subsection{Cooling techniques employed}
The other second generation solution of energy efficient data center is the use of air economizers in data centers.An air-side economizer brings outside air into a building and distributes it to the servers. This can lead to a significant savings in cooling energy of the data center.

We will study each of these factor individually and look at their impact on the cost. Based on this, we will devise an appropriate optimization technique which will minimize the energy cost while obeying all the constraints.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Expected Milestones}

 \subsection{Study real-time and synthetic workloads} We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011. This process is challenging because task-IDs, job-IDs are obfuscated for security reasons. Resources utilization such as CPU and memory has been normalized with the maximum resource usage. We are also studying synthetic workloads that are generated by sampling historical MapReduce cluster traces. Tentatively we would complete this by 10\textsuperscript{th}March,2015.
 \subsection{Data Center profile}In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-replicated data centers.We would complete this by 20\textsuperscript{th}March,2015.
\subsection{Feature Extraction} This phase will be identifying all the necessary features that contribute to the energy cost. These features will be used in the optimization of the energy cost. We aim to complete this by 5\textsuperscript{th}April,2015.
 \subsection{Implementing Optimization technique} This is the most critical component and expected to consume major part of project. We will be building optimization techniques which is adaptive to workload characteristics and data center's resource capacity and electricity costs.We aim to complete this by 25\textsuperscript{th}April,2015.
\subsection{Evaluation} This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above.We aim to complete this by 7\textsuperscript{th}May,2015.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{report}   %>>>> bibliography data in report.bib
\bibliographystyle{spiebib}   %>>>> makes bibtex use spiebib.bst

\end{document}

### 2015-03-14 doc/report/survey/article.tex
google workload code + project files

@@
 
- \subsection{Study real-time and synthetic workloads} We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.
+ \subsection{Study real-world and synthetic workloads} We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.
 This process is challenging because task-IDs, job-IDs are obfuscated for security reasons.

### 2015-03-14 src/GoogleTrace/JobEvent.cpp

@@ -1,0 +1,105 @@
+/*
+ * JobEvent.cpp
+ *
+ *  Created on: Mar 14, 2015
+ *      Author: Anirudh Jayakuamr
+ */
+#include "JobEvent.h"
+#include "common.h"
+#include <fstream>
+#include <string>
+#include <iostream>
+#include <stdlib.h>
+#include "GoogleTrace.h"
+#include <unordered_map>
+using namespace std;
+
+JobEvent::JobEvent()
+{
+}
+
+int JobEvent::Initialize(const std::string &filePath)
+{
+	ifstream fileStream(filePath);
+	string sLine;
+	vector<string> vTokens;
+	nRecords = 0;
+	while(!fileStream.eof())
+	{
+		fileStream >> sLine;
+		vTokens.clear();
+		Utils::split(sLine,',',vTokens);
+		char *pEnd;
+		vTimeStamp.push_back(strtoll(vTokens[TIMESTAMP].c_str(),&pEnd,10));
+		vJobID.push_back(strtoll(vTokens[JOBID].c_str(),&pEnd,10));
+		vEventType.push_back(strtol(vTokens[EVENT].c_str(),&pEnd,10));
+		vUserName.push_back(vTokens[USER]);
+		vSchedClass.push_back(strtol(vTokens[SCHEDCLASS].c_str(),&pEnd,10));
+		nRecords++;
+	}
+    cout << "Complete Initialization. Total records read:" << nRecords << endl;
+	fileStream.close();
+	return SUCCESS;
+}
+
+int JobEvent::GenerateCSVWithEventRecordsInSingleRow(const std::string &
+		fileWithJobIDs, const std::string &destFilePath)
+{
+	//load job ids in hash map
+	unordered_map<INT64_,Workload::Job> jobIDtoInfo;
+	string sLine;
+	char *pEnd; //dummy for API completeness
+	cout << "Loading JobIDs into hash table:" << endl;
+	ifstream srcFileStream(fileWithJobIDs);
+	while(!srcFileStream.eof())
+	{
+		srcFileStream >> sLine;
+		INT64_ nJobId = strtoll(sLine.c_str(),&pEnd,10);
+		auto it = jobIDtoInfo.find(nJobId);
+		if( it != jobIDtoInfo.end() ) {
+		    cout << "Duplicate Job ID:" << sLine << endl;
+		}
+		else {
[…]

### 2015-03-14 src/GoogleTrace/JobEvent.h

@@ -1,0 +1,51 @@
+/*
+ * JobEvent.h
+ *
+ *  Created on: Mar 12, 2015
+ *      Author: Anirudh Jayakumar
+ */
+
+#ifndef JOBEVENT_H_
+#define JOBEVENT_H_
+
+#include <string>
+#include <vector>
+#include "common.h"
+
+#define TIMESTAMP 	0
+#define JOBID     	2
+#define EVENT     	3
+#define USER      	4
+#define SCHEDCLASS  5
+class JobEvent
+{
+private:
+	/*
+	 * The job events table contains the following fields:
+		1. timestamp 2. missinginfo 3. jobID 4. eventtype
+		5. username 6. schedulingclass 7. jobname 8. logicaljobname
+
+		We only consider a subset of the above fields
+	 */
+	std::vector<INT64_> 		vTimeStamp;
+	std::vector<INT64_> 		vJobID;
+	std::vector<int> 			vEventType;
+	std::vector<std::string>	vUserName;
+	std::vector<int>       		vSchedClass;
+	int 						nRecords;
+public:
+	JobEvent();
+	int Initialize(const std::string &filePath);
+	//this method will create a partial trace, the other part of the trace
+	//file will be extracted from task events and SLA generators
+    int GenerateCSVWithEventRecordsInSingleRow(const std::string &fileWithJobIDs,
+    		const std::string &destFilePath);
+    /* add methods in future for newer requirements */
+private:
+
+
+};
+
+
+
+#endif /* JOBEVENT_H_ */

### 2015-03-14 src/GoogleTrace/TaskEvent.h

@@ -1,0 +1,15 @@
+/*
+ * TaskEvent.h
+ *
+ *  Created on: Mar 12, 2015
+ *      Author: Anirudh Jayakumar
+ */
+
+#ifndef TASKEVENT_H_
+#define TASKEVENT_H_
+
+
+
+
+
+#endif /* TASKEVENT_H_ */

### 2015-03-14 src/GoogleTrace/main.cpp

@@ -1,0 +1,14 @@
+/*
+ * main.cpp
+ *
+ *  Created on: Mar 14, 2015
+ *      Author: Anirudh Jayakumar
+ */
+
+
+
+
+int main(int argc,char *argv[])
+{
+	return 0;
+}

### 2015-03-14 src/includes/GoogleTrace.h

@@ -1,0 +1,83 @@
+#ifndef GOOGLETRACE_H_
+#define GOOGLETRACE_H_
+
+namespace Workload
+{
+
+
+struct Job
+{
+	INT64_ 			nJobID;  // job event
+	INT64_ 			nArrivalTime; // job event
+	INT64_  		nScheduledTime;  // job event
+	INT64_ 			nFinishTime;   // job event
+	int			    nSchedClass; // job event, may not need it once SLA is decided
+	int 			nTasks;
+	float			fTotalCPU;
+	float 			fTotalMem;
+	std::string 	sUser;      // job event
+	float 			fSLAParam1; // synthetic
+	float 			fSLAParam2; // synthetic
+	float 			fSLAParam3; // synthetic
+
+	Job()
+	{
+		nJobID 			= -1;
+		nArrivalTime 	= -1;
+		nScheduledTime  = -1;
+		nFinishTime 	= -1;
+		nSchedClass     = -1;
+		nTasks 			= 0;
+		fTotalCPU 		= 0;
+		fTotalMem 		= 0;
+		sUser 			= "";
+		fSLAParam1 		= 0;
+		fSLAParam2 		= 0;
+		fSLAParam3 		= 0;
+	}
+
+	INT64_ WaitTime()
+	{
+		return (nScheduledTime - nArrivalTime);
+	}
+
+	INT64_ RunningTime()
+	{
+		return (nFinishTime - nScheduledTime);
+	}
+
+	std::string GenerateCSV()
+	{
+		std::string retStr = std::to_string(nJobID) + ",";
+		retStr += std::to_string(nArrivalTime) + ",";
+		retStr += std::to_string(nScheduledTime) + ",";
+		retStr += std::to_string(nFinishTime) + ",";
+		retStr += std::to_string(nSchedClass) + ",";
+		retStr += std::to_string(nTasks) + ",";
+		retStr += std::to_string(fTotalCPU) + ",";
+		retStr += std::to_string(fTotalMem) + ",";
+		retStr += sUser + ",";
+		retStr += std::to_string(fSLAParam1) + ",";
+		retStr += std::to_string(fSLAParam2) + ",";
+		retStr += std::to_string(fSLAParam3);
+		return retStr;
+
+	}
+
+};
+
+class GoogleTrace
+{
+private:
+	std::vector<Job> vJobs;
+public:
+	int Initialize(const std::string &filePath);
[…]

### 2015-03-14 src/includes/common.h

@@ -1,0 +1,33 @@
+#ifndef COMMON_H_
+#define COMMON_H_
+
+#include <string>
+#include <vector>
+#include <sstream>
+typedef long long int INT64_;
+
+
+#define SUCCESS 0
+#define FAIL 1
+
+namespace Utils
+{
+
+
+	std::vector<std::string> &split(const std::string &s, char delim, std::vector<std::string> &elems) {
+	    std::stringstream ss(s);
+	    std::string item;
+	    while (std::getline(ss, item, delim)) {
+	        elems.push_back(item);
+	    }
+	    return elems;
+	}
+
+
+};
+
+
+
+#endif /* COMMON_H_ */
+
+

### 2015-03-27 src/GoogleTrace/GoogleTrace.cpp
task event parser + first complete compiling version

@@ -1,0 +1,132 @@
+/*
+ * GoogleTrace.cpp
+ *
+ *  Created on: Mar 26, 2015
+ *      Author: Anirudh Jayakumar
+ */
+
+
+#include "GoogleTrace.h"
+#include <string>
+#include <fstream>
+using namespace std;
+
+
+
+Workload::Job::Job()
+{
+	nJobID 			= -1;
+	nArrivalTime 	= -1;
+	nScheduledTime  = -1;
+	nFinishTime 	= -1;
+	nSchedClass     = -1;
+	nTasks 			= 0;
+	fTotalCPU 		= 0;
+	fTotalMem 		= 0;
+	sUser 			= "";
+	fSLAParam1 		= 0;
+	fSLAParam2 		= 0;
+	fSLAParam3 		= 0;
+	nCPUMissing     = 0;
+	nMemMissing     = 0;
+}
+
+int Workload::Job::Init(const std::string &sCSV)
+{
+	vector<string> vTokens;
+	Utils::Tokenize(sCSV,',',vTokens);
+	char *pEnd;
+	if(!vTokens[JOB_ID].empty())
+		nJobID = strtoll(vTokens[JOB_ID].c_str(),&pEnd,10);
+	if(!vTokens[JOB_ARRIVAL].empty())
+		nArrivalTime = strtoll(vTokens[JOB_ARRIVAL].c_str(),&pEnd,10);
+	if(!vTokens[JOB_SCHEDTIME].empty())
+		nScheduledTime = strtoll(vTokens[JOB_SCHEDTIME].c_str(),&pEnd,10);
+	if(!vTokens[JOB_FINISH].empty())
+		nFinishTime = strtoll(vTokens[JOB_FINISH].c_str(),&pEnd,10);
+	if(!vTokens[JOB_SCHEDCLASS].empty())
+		nSchedClass = strtol(vTokens[JOB_SCHEDCLASS].c_str(),&pEnd,10);
+	if(!vTokens[JOB_TASKS].empty())
+		nTasks = strtol(vTokens[JOB_TASKS].c_str(),&pEnd,10);
+	if(!vTokens[JOB_CPU].empty())
+		fTotalCPU = strtod(vTokens[JOB_CPU].c_str(),&pEnd);
+	if(!vTokens[JOB_MEM].empty())
+		fTotalMem = strtod(vTokens[JOB_MEM].c_str(),&pEnd);
[…]

### 2015-03-27 src/GoogleTrace/JobEvent.cpp

@@ -28,7 +28,7 @@ int JobEvent::Initialize(const std::string &filePath)
 	{
 		fileStream >> sLine;
 		vTokens.clear();
-		Utils::split(sLine,',',vTokens);
+		Utils::Tokenize(sLine,',',vTokens);
 		char *pEnd;
 		vTimeStamp.push_back(strtoll(vTokens[TIMESTAMP].c_str(),&pEnd,10));
 		vJobID.push_back(strtoll(vTokens[JOBID].c_str(),&pEnd,10));

### 2015-03-27 src/GoogleTrace/TaskEvent.cpp

@@ -1,0 +1,113 @@
+/*
+ * TaskEvent.cpp
+ *
+ *  Created on: Mar 25, 2015
+ *      Author: Anirudh Jayakumar
+ */
+
+#include "TaskEvent.h"
+#include "common.h"
+#include <iostream>
+#include <fstream>
+#include <set>
+#include <string>
+#include <utility>
+#include <algorithm>
+#include <unordered_map>
+#include <stdlib.h>
+#include "GoogleTrace.h"
+using namespace std;
+
+int FillEventInfoIntoCustomTrace(const std::string &sUniqueJobIdFile,const std::string &sEventsPath,
+		const std::string &sCustomTraceSource,const std::string &sCustomTraceDest)
+{
+	// load uniquejobIds
+	ifstream jobIDstream(sUniqueJobIdFile);
+	string sLine;
+	char *pEnd; //dummy for API completeness
+	set<INT64_> uniqueJobIds;
+	while(!jobIDstream.eof())
+	{
+		jobIDstream >> sLine;
+		INT64_ nJobId = strtoll(sLine.c_str(),&pEnd,10);
+		uniqueJobIds.insert(nJobId);
+	}
+	unordered_map<INT64_,JobEventsInfo> jobIDtoInfo;
+	// open each file
+	for (int fileIndex = 1; fileIndex <= 10; ++fileIndex)
+	{
+		// go through the traces
+		string sTraceName = sEventsPath + "/" + to_string(fileIndex) + ".csv";
+		cout << "Opening file" << sTraceName << endl;
+		ifstream eventTrace(sTraceName);
+		while(!eventTrace.eof())
+		{
+			eventTrace >> sLine;
+			vector<string> vTokens;
+			Utils::Tokenize(sLine,',',vTokens);
+			INT64_ nJobID = strtoll(vTokens[JOBID].c_str(),&pEnd,10);
+			// consider only job ids that we want
+			if(uniqueJobIds.find(nJobID) != uniqueJobIds.end() )
+			{
+				//see if the job already present in hash table
+				if( jobIDtoInfo.find(nJobID) == jobIDtoInfo.end())
+					jobIDtoInfo.insert(make_pair(nJobID,JobEventsInfo()));
+				// see if the event was already considered
+				int nTaskIndex = strtol(vTokens[TASKID].c_str(),&pEnd,10);
[…]

### 2015-03-27 src/GoogleTrace/TaskEvent.h

@@ -7,9 +7,41 @@
 
 #ifndef TASKEVENT_H_
 #define TASKEVENT_H_
+#include <map>
+#include <vector>
+#include <string>
+/*
+1. timestamp 2. missinginfo 3. jobID
+4. taskindex-withinthejob 5. machineID
+6. eventtype  7. username 8. schedulingclass 9. priority
+10. resource request for CPU cores 11. resource request for RAM
+12. resource request for local disk space 13. differentt-machine constraint
+*/
+
+#define JOBID 	2
+#define TASKID 	3
+#define CPU 	9
+#define MEM		10
+
+/* bookkeeping struct */
+
+struct JobEventsInfo
+{
+	std::vector<int> 	vEventIndices;
+	double 				dTotalCPU;
+	double 				dTotalMem;
+	int 				nCPUMissing;
+	int 				nMemMissing;
+	JobEventsInfo()
+	{
+		dTotalCPU = dTotalMem = 0.0;
+		nCPUMissing = nMemMissing = 0;
+	}
+};
 
 
-
-
+/* memory intensive function */
+int FillEventInfoIntoCustomTrace(const std::string &sUniqueJobIdFile,const std::string &sEventsPath,
+		const std::string &sCustomTraceSource,const std::string &sCustomTraceDest);
 
 #endif /* TASKEVENT_H_ */

### 2015-03-27 src/GoogleTrace/common.cpp

@@ -1,0 +1,18 @@
+/*
+ * common.cpp
+ *
+ *  Created on: Mar 25, 2015
+ *      Author: Anirudh Jayakumar
+ */
+
+#include "common.h"
+
+std::vector<std::string> &Utils::Tokenize(const std::string &s, char delim, std::vector<std::string> &elems) {
+	std::stringstream ss(s);
+	std::string item;
+	while (std::getline(ss, item, delim)) {
+		elems.push_back(item);
+	}
+	return elems;
+}
+

### 2015-03-27 src/GoogleTrace/main.cpp

@@ -5,10 +5,44 @@
  *      Author: Anirudh Jayakumar
  */
 
+#include <string>
+#include <iostream>
+#include "JobEvent.h"
+#include "TaskEvent.h"
+#include "common.h"
+using namespace std;
 
+const string sJobTrace("/Users/anirudhnair/Downloads/googletrace/clusterdata-2011-2/job_events/full.csv");
+const string sCompleteJobIDs("/Users/anirudhnair/Downloads/googletrace/clusterdata-2011-2/job_events/complete_jobIds");
+const string sCustomFile("/Users/anirudhnair/Downloads/googletrace/clusterdata-2011-2/job_events/custom.csv");
+const string sTaskTracePath("/Users/anirudhnair/Downloads/googletrace/clusterdata-2011-2/task_events");
 
+int GenerateJobEventInfo()
+{
+	JobEvent oJobTrace;
+	if( oJobTrace.Initialize(sJobTrace) == FAIL)
+	{
+		cout << "Jobtrace init fail" << endl;
+		return FAIL;
+	}
+
+	if( oJobTrace.GenerateCSVWithEventRecordsInSingleRow(sCompleteJobIDs,sCustomFile) == FAIL)
+	{
+		cout << "GenerateCSVWithEventRecordsInSingleRow failed" << endl;
+		return FAIL;
+	}
+	return SUCCESS;
+}
+
+int GenerateTaskEventInfo()
+{
+	string sDestFile = "/Users/anirudhnair/Downloads/googletrace/clusterdata-2011-2/job_events/custom_dest.csv";
+	return FillEventInfoIntoCustomTrace(sCompleteJobIDs,sTaskTracePath,sCustomFile,sDestFile);
+}
 
 int main(int argc,char *argv[])
 {
+
+	GenerateJobEventInfo();
 	return 0;
 }

### 2015-03-27 src/includes/GoogleTrace.h

@@ -1,68 +1,51 @@
 #ifndef GOOGLETRACE_H_
 #define GOOGLETRACE_H_
 
+#include "common.h"
+
+#define JOB_ID  		0
+#define JOB_ARRIVAL  	1
+#define JOB_SCHEDTIME  	2
+#define JOB_FINISH  	3
+#define JOB_SCHEDCLASS 	4
+#define JOB_TASKS	  	5
+#define JOB_CPU		  	6
+#define JOB_MEM		  	7
+#define JOB_USER	  	8
+#define JOB_SLA1	  	9
+#define JOB_SLA2	  	10
+#define JOB_SLA3	  	11
+
 namespace Workload
 {
 
 
 struct Job
 {
+	// any changes to the variables here must be updated in the
+	// trace format xml
 	INT64_ 			nJobID;  // job event
 	INT64_ 			nArrivalTime; // job event
 	INT64_  		nScheduledTime;  // job event
 	INT64_ 			nFinishTime;   // job event
 	int			    nSchedClass; // job event, may not need it once SLA is decided
 	int 			nTasks;
-	float			fTotalCPU;
-	float 			fTotalMem;
+	double			fTotalCPU;
+	double 			fTotalMem;
 	std::string 	sUser;      // job event
-	float 			fSLAParam1; // synthetic
-	float 			fSLAParam2; // synthetic
-	float 			fSLAParam3; // synthetic
+	double 			fSLAParam1; // synthetic
+	double 			fSLAParam2; // synthetic
+	double 			fSLAParam3; // synthetic
 
-	Job()
-	{
-		nJobID 			= -1;
-		nArrivalTime 	= -1;
-		nScheduledTime  = -1;
-		nFinishTime 	= -1;
-		nSchedClass     = -1;
-		nTasks 			= 0;
-		fTotalCPU 		= 0;
-		fTotalMem 		= 0;
-		sUser 			= "";
-		fSLAParam1 		= 0;
-		fSLAParam2 		= 0;
-		fSLAParam3 		= 0;
-	}
+	Job();
+	int 			Init(const std::string &sCSV);
+	INT64_ 			WaitTime();
+	INT64_ 			RunningTime();
+	std::string 	GenerateCSV();
 
-	INT64_ WaitTime()
-	{
-		return (nScheduledTime - nArrivalTime);
-	}
-
-	INT64_ RunningTime()
-	{
-		return (nFinishTime - nScheduledTime);
-	}
-
[…]

### 2015-03-27 src/includes/common.h

@@ -13,16 +13,7 @@ typedef long long int INT64_;
 namespace Utils
 {
 
-
-	std::vector<std::string> &split(const std::string &s, char delim, std::vector<std::string> &elems) {
-	    std::stringstream ss(s);
-	    std::string item;
-	    while (std::getline(ss, item, delim)) {
-	        elems.push_back(item);
-	    }
-	    return elems;
-	}
-
+	std::vector<std::string> &Tokenize(const std::string &s, char delim, std::vector<std::string> &elems);
 
 };
 

### 2015-03-30 doc/report/survey/article.tex
changes to survey to make tables and also list

--- 2015-03-02 doc/report/survey/article.tex
@@
 }
+
+\newenvironment{list1}{
+  \begin{list}{\ding{113}}{%
+      \setlength{\itemsep}{0in}
+      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
+      \setlength{\topsep}{0in} \setlength{\partopsep}{0in}
+      \setlength{\leftmargin}{0.17in}}}{\end{list}}
+\newenvironment{list2}{
+  \begin{list}{$\bullet$}{%
+      \setlength{\itemsep}{0in}
+      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
+      \setlength{\topsep}{0in} \setlength{\partopsep}{0in}
+      \setlength{\leftmargin}{0.2in}}}{\end{list}}
 
@@
 
-\subsection{Electricity pricing} 
-We use the day-ahead hourly electricity prices to get the electricity cost of the data center which will enable us to provision workloads efficiently.
-Moreover, the prices in the day-ahead market fluctuate smoothly and can be predicted using time series forecasting techniques.
-\subsection{Energy source in data centers}
-To meet the energy requirements of data centers there is increasing adoption of renewable energy supply techniques (solar, wind).
+\begin{list2}
+\item \textbf{Electricity pricing} We use the day-ahead hourly electricity prices to get the electricity cost of the data center which will enable us to provision workloads efficiently.\
+\item \textbf{Energy source in data centers} To meet the energy requirements of data centers there is increasing adoption of renewable energy supply techniques (solar, wind).
 This has resulted in data centers being powered at least partially with green energy.
-We wish to effectively use the green energy to reduce costs.
-\subsection{Cooling techniques employed}
-The other second generation solution of energy efficient data center is the use of air economizers in data centers.An air-side economizer brings outside air into a building and distributes it to the servers.
-This can lead to a significant savings in cooling energy of the data center.
-
-We will study each of these factor individually and look at their impact on the cost.
-Based on this, we will devise an appropriate optimization technique which will minimize the energy cost while obeying all the constraints.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\
+\item \textbf{Cooling techniques employed} The other second generation solution of energy efficient data center is the use of air economizers in data centers.
+An air-side economizer brings outside air into a building and distributes it to the servers.
+\
+\end{list2}
 
@@
 
- \subsection{Study real-time and synthetic workloads} We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.
-This process is challenging because task-IDs, job-IDs are obfuscated for security reasons.
-Resources utilization such as CPU and memory has been normalized with the maximum resource usage.
-We are also studying synthetic workloads that are generated by sampling historical MapReduce cluster traces.
-Tentatively we would complete this by 10\textsuperscript{th}March,2015.
- \subsection{Data Center profile}In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-replicated data centers.We would complete this by 20\textsuperscript{th}March,2015.
-\subsection{Feature Extraction} This phase will be identifying all the necessary features that contribute to the energy cost.
-These features will be used in the optimization of the energy cost.
-We aim to complete this by 5\textsuperscript{th}April,2015.
- \subsection{Implementing Optimization technique} This is the most critical component and expected to consume major part of project.
-We will be building optimization techniques which is adaptive to workload characteristics and data center's resource capacity and electricity costs.We aim to complete this by 25\textsuperscript{th}April,2015.
-\subsection{Evaluation} This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above.We aim to complete this by 7\textsuperscript{th}May,2015.
+    \begin{tabular}{ | p{3cm} |l  |p{12cm}|  }
+    \hline
+    \bf{Item} & \bf{Deadline} & \bf{Description}   \\ \hline
+    Study real-world and synthetic workloads & 10\textsuperscript{th} March,2015  & We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.
+\\ \hline
+    Data Center profile & 20\textsuperscript{th} March,2015  & In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-distributed data centers.
+\\ \hline
+    Feature Extraction & 25\textsuperscript{th} April,2015  & This phase will be identifying all the necessary features that contribute to the energy cost.
+These features will be used in the optimization of the energy cost \\ \hline
+    Implementing Optimization technique & 5\textsuperscript{th} April,2015  & This is the most critical component and expected to consume major part of project.
+We will be building optimization techniques which is adaptive to workload characteristics and data center's resource capacity and electricity costs \\ \hline
+    Evaluation & 7\textsuperscript{th} May,2015  & This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above \\ \hline
+
+    \end{tabular}
 

### 2015-04-01 doc/report/midterm/article.tex
mid term report outline

--- 2015-03-30 doc/report/survey/article.tex
@@
 
-\section{Approaches }
+\section{Cloud Workload }
 
-We wish to consider all the major factors that impact the energy costs at the data center. Some of the factors and its impact are discussed in the below sub-section.
+\subsection{Workload Traces}
 
-\begin{list2}
-\item \textbf{Electricity pricing} We use the day-ahead hourly electricity prices to get the electricity cost of the data center which will enable us to provision workloads efficiently.\
-\item \textbf{Energy source in data centers} To meet the energy requirements of data centers there is increasing adoption of renewable energy supply techniques (solar, wind).
-This has resulted in data centers being powered at least partially with green energy.
-\
-\item \textbf{Cooling techniques employed} The other second generation solution of energy efficient data center is the use of air economizers in data centers.
-An air-side economizer brings outside air into a building and distributes it to the servers.
-\
-\end{list2}
+\subsection{Google Workload Trace}
 
-\section{Expected Milestones}
+\subsection{Custom Trace Format}
 
-    \begin{tabular}{ | p{3cm} |l  |p{12cm}|  }
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Data Center Profiling }
+
+Provide some introduction about out profiling stratergy
+
+\subsection{Geo-Distributed Data Center}
+
+\subsection{Profile Parameters}
+\subsubsection{Temperature}
+\subsubsection{Electricity Pricing}
+\subsubsection{Energy Source}
+
+\subsection{Air Economizer}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Scheduler simulator}
+
+Provide basic idea about our strategy
+
+\subsection{Architecture}
+Introduce cloudsim
+\subsubsection{Components}
+\subsubsection{Integration of traces and data center profile}
+\subsubsection{Design}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Scheduling Algorithm}
+
+Provide basic idea about our strategy
+
+\subsection{Features considered}
+\subsubsection{Energy Cost Model}
+%\subsubsection{Optimization Techiques}
+%\subsubsection{Design}
+
+\section{ Milestones}
+
+    \begin{tabular}{ | p{3cm} |l  |p{11cm}|  }
     \hline
@@
 \\ \hline
-    Feature Extraction & 25\textsuperscript{th} April,2015  & This phase will be identifying all the necessary features that contribute to the energy cost.
+    Simulator design & 5\textsuperscript{th} April,2015  & Implement the simulator infrastructure for multi-datacenter environment.  \\ \hline
+    Feature Extraction & 8\textsuperscript{th} April,2015  & This phase will be identifying all the necessary features that contribute to the energy cost.
 These features will be used in the optimization of the energy cost \\ \hline
-    Implementing Optimization technique & 5\textsuperscript{th} April,2015  & This is the most critical component and expected to consume major part of project.
+    Implementing Optimization technique & 25\textsuperscript{th} April,2015  & This is the most critical component and expected to consume major part of project.
 We will be building optimization techniques which is adaptive to workload characteristics and data center's resource capacity and electricity costs \\ \hline

### 2015-04-04 src/GoogleTrace/plots/ArrivalRates.py
scaled up trace + plot code + plots

@@ -1,0 +1,50 @@
+import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
+import os
+
+f = open("/Users/anirudhnair/code/GeoSched/workloads/google/5_1arrival.csv")
+a1 = f.readlines()
+f.close()
+f = open("/Users/anirudhnair/code/GeoSched/workloads/google/5_2arrival.csv")
+a2 = f.readlines()
+f.close()
+f = open("/Users/anirudhnair/code/GeoSched/workloads/google/5_3arrival.csv")
+a3 = f.readlines()
+f.close()
+f = open("/Users/anirudhnair/code/GeoSched/workloads/google/5_4arrival.csv")
+a4 = f.readlines()
+f.close()
+f = open("/Users/anirudhnair/code/GeoSched/workloads/google/5_5arrival.csv")
+a5 = f.readlines()
+f.close()
+
+plt.rc('text', usetex=False)
+plt.rc('font', family='serif')
+
+#xLabel_ = [0,15,30,45,60,75,90,105,120]
+#plt.xticks(range(len(xLabel_)), xLabel_, size='medium')
+
+#plt.scatter(range(len(a1)),a1)
+plt.plot(range(len(a1)), a1, linestyle='-' ,marker='o', markersize=4,label='Trace A')
+
+#plt.scatter(range(len(a2)),a2)
+plt.plot(range(len(a2)), a2, linestyle='-' , marker='^', markersize=4,label='Trace B')
+
+#plt.scatter(range(len(a3)),a3)
+plt.plot(range(len(a3)), a3, linestyle='-' , marker='+', markersize=4,label='Trace C')
+
+#plt.scatter(range(len(a4)),a4)
+plt.plot(range(len(a4)), a4, linestyle='-' , marker='s', markersize=4,label='Trace D')
+
+#plt.scatter(range(len(a5)),a5)
+plt.plot(range(len(a5)), a5,  linestyle='-' , marker='x', markersize=4,label='Trace E')
+
+plt.ylim([200,1400])
+legend = plt.legend(loc='upper right', shadow=True)
+
+plt.ylabel(r'Arriving Job Count')
+plt.xlabel(r'Hours')
+
+
+plt.savefig('arrivalrate.jpg', format='jpg', dpi=600)
+plt.clf()

### 2015-04-04 src/GoogleTrace/plots/SchedClassPie.py

@@ -1,0 +1,89 @@
+"""
+Make a pie chart - see
+http://matplotlib.sf.net/matplotlib.pylab.html#-pie for the docstring.
+
+This example shows a basic pie chart with labels optional features,
+like autolabeling the percentage, offsetting a slice with "explode",
+adding a shadow, and changing the starting angle.
+
+"""
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
+from matplotlib import font_manager as fm
+# make a square figure and axes
+#plt.figure(1, figsize=(5,2.5))    
+#ax = plt.axes([0.1, 0.1, 0.8, 0.8])
+
+# The slices will be ordered and plotted counter-clockwise.
+labels = r"C0", r"C1", r"C2", r"C3"
+
+mpl.rcParams['font.size'] = 9.0
+mpl.rcParams['lines.linewidth'] = 0.1
+color_ = ['r', 'y', 'g', 'b']
+plt.rc('font', family='serif')
+explode = (0, 0, 0, 0)
+the_grid = GridSpec(1, 5)
+
+plt.axis('equal')
+fracs = [30.43,45.36,24.19,0.02]
+plt.subplot(the_grid[0, 0], aspect=1)
+patches, texts, autotexts = plt.pie(fracs, explode=explode,labels=labels,labeldistance=1.05,
+                autopct="%1.2f%%", shadow=False, startangle=0, colors=color_,radius=1.2)
+for w in patches:
+    w.set_linewidth( 0 )
+proptease = fm.FontProperties()
+proptease.set_size('x-small')
+plt.setp(autotexts, fontproperties=proptease)
+plt.setp(texts, fontproperties=proptease)
+plt.title(r"Trace A")
+fracs = [30.97,44.44,24.57,0.02]
+plt.subplot(the_grid[0, 1], aspect=1)
+patches, texts, autotexts = plt.pie(fracs, explode=explode,labels=labels,labeldistance=1.05,
+                autopct="%1.2f%%", shadow=False, startangle=0, colors=color_,radius=1.2)
+for w in patches:
+    w.set_linewidth( 0 )
+proptease = fm.FontProperties()
+proptease.set_size('x-small')
+plt.setp(autotexts, fontproperties=proptease)
+plt.setp(texts, fontproperties=proptease)
+
+plt.title(r"Trace B")
+fracs = [31.93,43.09,24.96,0.03]
[…]

### 2015-04-04 src/GoogleTrace/plots/resourceCDF.py

@@ -1,0 +1,64 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.stats import cumfreq
+import scikits.statsmodels as sm
+from matplotlib import pyplot
+from matplotlib.gridspec import GridSpec
+
+ 
+def plot_func_cpu(filename,index_,marker):
+    f = open(filename)
+    a1 = f.readlines()
+    f.close()
+    runtimes = [float(val) for val in a1]
+    runtimes = np.array(runtimes) # my array of numbers
+    sorted_data = np.sort(runtimes)
+    max_val = np.max(sorted_data)
+    min_val = np.min(sorted_data)
+    print max_val, "," , min_val
+    hi,label_ = np.histogram(sorted_data, bins=np.arange(0,max_val + 0.01,0.01))
+    csum = np.cumsum(hi) 
+    line_id = 'Trace ' + index_
+    plt.plot(label_, np.insert(csum, 0, 0),label=line_id,marker=marker,markersize=4)
+   
+def plot_func_mem(filename,index_,marker):
+    f = open(filename)
+    a1 = f.readlines()
+    f.close()
+    runtimes = [float(val) for val in a1]
+    runtimes = np.array(runtimes) # my array of numbers
+    sorted_data = np.sort(runtimes)
+    max_val = np.max(sorted_data)
+    min_val = np.min(sorted_data)
+    print max_val, "," , min_val
+    hi,label_ = np.histogram(sorted_data, bins=np.arange(0,max_val + 0.0005,0.0005))
+    csum = np.cumsum(hi) 
+    line_id = 'Trace ' + index_
+    plt.plot(label_, np.insert(csum, 0, 0),label=line_id,marker=marker,markersize=4)
+
+
+plot_func_cpu('/Users/anirudhnair/code/GeoSched/workloads/google/5_1cpu','A','o')
+plot_func_cpu('/Users/anirudhnair/code/GeoSched/workloads/google/5_2cpu','B','^')
+plot_func_cpu('/Users/anirudhnair/code/GeoSched/workloads/google/5_3cpu','C','+')
+plot_func_cpu('/Users/anirudhnair/code/GeoSched/workloads/google/5_4cpu','D','s')
+plot_func_cpu('/Users/anirudhnair/code/GeoSched/workloads/google/5_5cpu','E','x')
+plt.ylabel(r'Job Count')
[…]

### 2015-04-04 src/GoogleTrace/plots/runtime.py

@@ -1,0 +1,38 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.stats import cumfreq
+import scikits.statsmodels as sm
+from matplotlib import pyplot
+ 
+def plot_func(filename,index_,marker):
+    f = open(filename)
+    a1 = f.readlines()
+    f.close()
+    runtimes = [int(val) for val in a1]
+    runtimes = np.array(runtimes) # my array of numbers
+    sorted_data = np.sort(runtimes)
+    sorted_data = np.divide(sorted_data,1000000)
+    max_time = np.max(sorted_data) 
+    max_time = max_time + (100 - max_time%100)
+    hi,label_ = np.histogram(sorted_data, bins=range(0,max_time + 100,100))
+    csum = np.cumsum(hi) 
+    line_id = 'Trace ' + index_
+    plt.plot(label_, np.insert(csum, 0, 0),label=line_id,marker=marker,markersize=4)
+   
+
+
+plot_func('/Users/anirudhnair/code/GeoSched/workloads/google/5_1runtime.csv','A','o')
+plot_func('/Users/anirudhnair/code/GeoSched/workloads/google/5_2runtime.csv','B','^')
+plot_func('/Users/anirudhnair/code/GeoSched/workloads/google/5_3runtime.csv','C','+')
+plot_func('/Users/anirudhnair/code/GeoSched/workloads/google/5_4runtime.csv','D','s')
+plot_func('/Users/anirudhnair/code/GeoSched/workloads/google/5_5runtime.csv','E','x')
+
+plt.xscale('log',basex=2)
+#plt.yscale('log',basey=2)
+legend = plt.legend(loc='lower right', shadow=True)
+plt.ylim([20000,80000])
+plt.ylabel(r'Job Count')
+plt.xlabel(r'Job duration(secs)')
+plt.savefig('runtime.jpg', format='jpg', dpi=600)
+plt.clf()
+

### 2015-04-05 doc/report/midterm/article.tex
tex + bib

@@
 \usepackage[sort, numbers]{natbib}
+\usepackage{hyperref}
+\usepackage{adjustbox,lipsum}
+\graphicspath{ {/Users/anirudhnair/code/GeoSched/src/GoogleTrace/plots} }
+\usepackage{subfig}
+
 \title{GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers} 
@@
 
+\iffalse
+
 Other research related to cooling efficiency include the work by Liu et all \cite{liu2012renewable}  that reduces electricity cost and environmental impact using a holistic approach that integrates renewable supply, dynamic pricing, and cooling supply including chiller and outside air cooling, with IT workload planning to improve the overall sustainability of data center operations.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\fi
 
 \section{Cloud Workload } 
+With the increase in the adoption of cloud computing by businesses around the world, more and more workloads are moving out of traditional data centers into the cloud.
+A cloud data center typically handles a range of IT workloads.
+These workloads include service-critical and latency sensitive interactive applications that run 24x7 such as instant messengers, e-mail applications and other internet services, and batch-style applications such as scientific HPC applications, BigData applications like financial analysis and MapReduce applications.
+A thoughtful management of these workloads are necessary not only to provide customers with low-latency services but also to reduce the energy consumption of the data centers that run these workloads.
+A good way to study these workloads are by analyzing real-world or synthetic workload traces.
+We look at workload traces in detail in section \ref{traces} and \ref{googletrace}
 
-\subsection{Workload Traces}
-
-\subsection{Google Workload Trace}
+\subsection{Workload Traces} \label{traces}
+Unfortunately, there is a lack of real-world traces in the cloud computing research community.
+As a result, there are limited work done to understand the diversity in cloud workloads.
+The first large-scale workload trace was released by Google in 2011\cite{googletracedata}, featuring traces from over 12,000 servers over a period of 29 days.
+Various studies have been carried out to understand the characteristic of the Google trace data.
+Other real-world traces include Wikipedia request trace \cite{urdaneta2009wikipedia} and Hadoop logs from Yahoo!
+\cite{yahootrace}There have also been attempts to generate synthetic workload traces using known real-world workload characteristics \cite{beitch2010rain} \cite{wang2011towards}.
+For the sake of this work, we have decided to use the workload trace from Google.
 
 \subsection{Custom Trace Format}
+To have a common trace format to represent different real-world and synthetic workload traces, we have formulated a trace format similar to standard workload format (swf) \cite{chapin1999benchmarks} defined for grid and supercomputing traces.
+Table \ref{tab:format} provides information about each field int the trace format.
+
+\begin {table}[ht]
+\caption {Trace format description} \label{tab:format} 
+ \scalebox{0.9}{\begin{tabular}{ | p{4cm} |l  |p{11 cm}|  }
+    \hline
+    \bf{Field Name} & \bf{Type} & \bf{Description}   \\ \hline
+     JobID & INTEGER  & Unique job identifier for each job\\ \hline
+     Arrival Time & INTEGER & Arrival time in microseconds \\ \hline
+     Scheduled Time & INTEGER  & Arrival time in microseconds\\ \hline
+     Finish Time & INTEGER  & Arrival time in microseconds\\ \hline
+     Schedule class & INTEGER  & Latency class of the job. \\ \hline
+     Tasks & INTEGER  & Total number of tasks in the job. \\ \hline
+     Total CPU& INTEGER  & Aggregate sum of CPU used by all the tasks in the job\\ \hline
+     Total Memory& INTEGER   & Aggregate sum of memory used by all the tasks in the job \\ \hline
+     User& INTEGER  & Unique user identifier \\ \hline
+     SLA Param1&  FLOAT & SLA parameter placeholder\\ \hline
+     SLA Param2&  FLOAT & SLA parameter placeholder\\ \hline
+     SLA Param3&  FLOAT & SLA parameter placeholder\\ \hline
+    \end{tabular}}
+\end{table}
+
+\subsection{Google Workload Trace} \label{googletrace}
+Google trace has job and task information in different files.
+We consolidated all the required information and generated a single work trace file with the format described in the previous section.
+For our work, we model five different data centers located at different parts of the globe.
+A detailed explanation about the data center profile are given in section \ref{sec:datacenter}.
+To simulate workload execution in these five data centers we have extracted five set of traces each worth five days of incoming workload from the original google trace.
+So, this covers 25 days of traces out of the 29 available days.
+We label these 5 set of traces from A to E.
+Additionally, in our experiments we wish to only consider those jobs that end without any failures.
+Therefore, we clean the original traces by removing all jobs that is either failed or get killed.
+
+\begin{figure}[ht] 
+\centering
+\includegraphics[scale=0.40]{arrivalrate.png}
+\caption{Arrival rate of jobs }
+\label{fig:arrivalrate}
+\end{figure}
+
+Now we will look at some characteristics of the traces.
+Table \ref{tab:tracestat} describes the statistics for each of the five traces grouped by the scheduling class of the job.
+In google traces, CPU and memory usage are normalized relative to the largest capacity of the resource on any machine in the trace (which is 1.0).
+Most of the class 0 and class 2 jobs are batch jobs with less latency requirements.
+Generally, these are the kind of jobs that can be provisioned on other data centers as their deadlines are not strict.
+You will find that as the sensitivity increases the running time of these jobs also increase.
+High-sensitive jobs are generally service jobs and run of longer periods of time.
+Figure \ref{fig:pie} shows a pie chart of the job counts in different schedule classes for all the 5 traces.
+You will notice that the job count for class 3 (high latency-sensitivity) jobs are low.
+This is because these jobs generally run for long period of time, typically months.
+Since we only consider jobs that start and end within the 29 day range such jobs are missed out in our traces.
+For our work, we will consider class-0 and class-1 jobs as candidates for optimizing the provisioning.
+We leave, class-2 and class-3 jobs to run on the data center where the job arrived since these are high-sensitive and their SLA demands will be strict.
+
+\begin{figure}[]
+\centering
+\includegraphics[scale=0.65]{pie}
+\caption{ Distribution of job count in various schedule classes}
+ \label{fig:pie}
+\end{figure}
+
+\begin{table}
+\caption {Trace set statistics} \label{tab:tracestat} 
+\adjustbox{max width=1\textwidth}{
+
+\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
+\hline
+ & \multicolumn{3}{l|}{\bf{Trace A}} & \multicolumn{3}{l|}{\bf{Trace B}} & \multicolumn{3}{l|}{\bf{Trace C}} & \multicolumn{3}{l|}{\bf{Trace D}} & \multicolumn{3}{l|}{\bf{Trace E}} \\ \hline
+\textbf{\begin{tabular}[c]{@{}l@{}}Sched\\ Class\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg \\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} \\ \hline
+0 & 0.075 & 0.051 & 415 & 0.123 & 0.096 & 345 & 0.083 & 0.065 & 385 & 0.093 & 0.092 & 373 & 0.093 & 0.074 & 333 \\ \hline
+1 & 0.087 & 0.029 & 729 & 0.086 & 0.032 & 760 & 0.082 & 0.028 & 711 & 0.073 & 0.030 & 743 & 0.093 & 0.043 & 693 \\ \hline
+2 & 0.067 & 0.043 & 1028 & 0.096 & 0.058 & 872 & 0.088 & 0.057 & 918 & 0.073 & 0.045 & 707 & 0.084 & 0.049 & 832 \\ \hline
+3 & 0.088 & 0.12 & 1105 & 0.099 & 0.12 & 82 & 0.101 & 0.012 & 7961 & 0.081 & 0.051 & 17135 & 0.089 & 0.061 & 9021 \\ \hline
+\end{tabular}
+}
+\end{table}
+
+Figure \ref{fig:arrivalrate} shows the variation in job arrival for the traces.
+As you can see there is both intra-trace variation as well as inter-trace variation.
+This is representative of the real-world cloud workload pattern where bursty arrivals are common.
+There are opportunities to provision jobs to other data centers if the traffic at a particular data center is high.
+What is also important to notice is that the scheduling decision should be made quickly when the arrival rates of jobs are high, else there are chances for these jobs to miss the SLA.
+Hence, even the scheduler should be scalable during high volume traffic.
+
+Figure \ref{fig:runtime} shows the CDF of the job runtime.
+As you can see most of the jobs have runtime less than 1024 seconds.
+This pattern is consistent for all the trace sets.
+There are a few jobs that have higher run-times.
+As explained earlier, these jobs are primarily internet-service jobs that run for months.
+
+\begin{figure}[] 
+\centering
+\includegraphics[scale=0.40]{runtime}
+\caption{CDF of the job duration}
+\label{fig:runtime}
+\end{figure}
+
+In google workload trace, it is tough to study the resource usage since these value are scaled down and therefore and accurate picture of the resource utilization is tough to generate.
+In figure \ref{fig:resource} we can see that most jobs request a small portion of the cluster resources.
+Most of the resources in the google cluster are used by the few service jobs that are present.
+The large number of batch jobs only use a small percentage of the resources \cite{mishra2010towards} \cite{reiss2012heterogeneity}.
+From the above statistics, we believe that google workload trace provides an accurate representation of the real-world cloud datacenter workload.
+
+\begin{figure} %
+    \centering
+    \subfloat[CPU resource ]{{\includegraphics[width=9cm]{cpu} }}%
+    \subfloat[RAM resource]{{\includegraphics[width=9cm]{mem} }}%
+    \caption{CDF of resource requested by jobs}%
+    \label{fig:resource}%
+\end{figure}
 
@@
 
-\section{Data Center Profiling }
+\section{Data Center Profiling } \label{sec:datacenter}
 
@@
 
-    \begin{tabular}{ | p{3cm} |l  |p{11cm}|  }
+    \begin{tabular}{ | p{4cm} |l  |p{11 cm}|  }
     \hline
     \bf{Item} & \bf{Deadline} & \bf{Description}   \\ \hline
-    Study real-world and synthetic workloads & 10\textsuperscript{th} March,2015  & We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.
+    Study real-world and synthetic workloads & 10\textsuperscript{th} March  & We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.
 \\ \hline
-    Data Center profile & 20\textsuperscript{th} March,2015  & In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-distributed data centers.
+    Data Center profile & 20\textsuperscript{th} March  & In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-distributed data centers.
 \\ \hline
-    Simulator design & 5\textsuperscript{th} April,2015  & Implement the simulator infrastructure for multi-datacenter environment.  \\ \hline
-    Feature Extraction & 8\textsuperscript{th} April,2015  & This phase will be identifying all the necessary features that contribute to the energy cost.
-These features will be used in the optimization of the energy cost \\ \hline
-    Implementing Optimization technique & 25\textsuperscript{th} April,2015  & This is the most critical component and expected to consume major part of project.
-We will be building optimization techniques which is adaptive to workload characteristics and data center's resource capacity and electricity costs \\ \hline
-    Evaluation & 7\textsuperscript{th} May,2015  & This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above \\ \hline
+    Simulator design and implementation& 12\textsuperscript{th} April  & Implement the simulator infrastructure for multi-datacenter environment.  \\ \hline
+    Feature Extraction and Optimization & 25\textsuperscript{th} April  & Here we identifying all the necessary features that contribute to the energy cost  and build an optimization technique which is adaptive to workload characteristics and data center's resource capacity and electricity costs \\ \hline
+    Evaluation & 7\textsuperscript{th} May  & This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above \\ \hline
 

### 2015-04-05 doc/report/midterm/article.tex
changed margins
Added DC Profiling

@@
 %%\documentclass[]{spie}  %>>> use for US letter paper
-\documentclass[12pt]{spie}  %>>> use this instead for A4 paper
+\documentclass[a4papert,12pt]{spie}  %>>> use this instead for A4 paper
 %%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
@@
 \usepackage{subfig}
+\usepackage[margin=0.5in]{geometry}
 
 \title{GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers} 
-\author{Anirudh Jayakumar and Harshit Dokania \\
-University of Illinois at Urbana-Champaign \\
-\{ajayaku2,hdokani2\}@illinois.edu 
-}
+%\author{Anirudh Jayakumar and Harshit Dokania \\
+%University of Illinois at Urbana-Champaign 
+%\{ajayaku2,hdokani2\}@illinois.edu 
+%}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Data Center Profiling } 
+Google data-centers show 50\% less energy consumption than a typical data center.
+Google is able to achieve a trailing twelve-month (TTM) PUE of 1.12 across all data centers, in all seasons, including all sources of overhead.
+These data-centers have varied characteristics over time, which will be discussed in detail in this section.
+\subsection{Geo-Distributed Data Center}
+We have chosen to simulate Google data centers that do the work for us around the world, around the clock at five different locations.
+These consists of two of them in United States namely Berkeley County, South Carolina and Lenoir, North Carolina.
+Other data centers that we have simulated are located one each in Europe, Asia, and South ?America.
+\subsection{Profile Parameters}
+There are five techniques that Google typically uses to minimize its energy consumption and electricity costs.
+Frequent measuring of Power Usage Effectiveness is the key for determining the power usage for non-computing functions like cooling and power distribution.
+Secondly they manage airflow at data-centers, which allows cooling the hot spots available in data-centers.
+Moreover by setting thermostat to 80$^{\circ}$F it saves significant amount of electricity consumption and costs.
+Google researchers propose maximum usage of ambient temperature through evaporative cooling, cold reservoir, and airside economizer.
+This discards conventional chillers, which are the source of maximum energy consumption at data centers.
+Altogether Google data continuously draw almost 260 million watts, which is about a quarter of the output of a nuclear power plant.
+Google uses renewable energy to power more than 35\% of its operations.
+These are wind, solar energy typically available at all data centers.
 
-\section{Data Center Profiling } \label{sec:datacenter}
+Google results show that data centers incur high PUE during summer and low PUE during winters.
+It clearly shows that surrounding temperature plays a critical role in deciding energy consumption at data centers.
+Data centers typically purchase wholesale electricity where the price is determined by day ahead market.
+\begin{figure} %
+    \centering
+    \subfloat[Daily Average Temperature ]{{\includegraphics[width=9cm]{AverageTemp} }}%
+    \subfloat[Daily Maximum Variation]{{\includegraphics[width=9cm]{Variation} }}%
+    \caption{Temperature variation for the year 2013}%
+    \label{fig:temp}%
+\end{figure}
 
-Provide some introduction about out profiling stratergy
+\subsubsection{Temperature}
+We use temperature as a key parameter to decide the location of job, as temperatures are diverse across data centers and encounters frequent fluctuations in temperature.
+In this case we make our assumption concrete by doing an empirical analysis of historical climate data, which provides substantial evidence.
+We extracted these historical climate data from various data repositories of the National Climate Data Center for all five locations covering the entire one-year period of 2013.
 
-\subsection{Geo-Distributed Data Center}
+Figure \ref{fig:temp} a plots the daily average temperature for five different locations in United States, Europe, South-America, and Asia.
+As expected we see high diversity in temperature at different locations.
+For example, North Carolina, Ireland seems to be equally favorable for cooling during the month of November {-} March.
+Chile and Singapore show almost no variation in temperature across the year.
+However, Chile is favorable for cooling conditions compared to other data centers in the month of May {-} July.
 
-\subsection{Profile Parameters}
-\subsubsection{Temperature}
+Figure \ref{fig:temp} b plots the daily maximum variation in temperature across the day i.e.
+difference of maximum temperature and minimum temperature in a day.
+It is worth noting that almost every location faces a variation of up to 40$^{\circ}$F during different periods.
+These highs and lows do not occur at the same time across different regions due to time differences.
+It is evident from the results that there is no single location that is always cooling efficient.
+So we conclude that we need a dynamic scheduler, which caters to the locations that have lower temperature and henceforth save overall energy costs.
+
 \subsubsection{Electricity Pricing}

### 2015-04-05 doc/report/midterm/article.tex
added more stuff to data center profiling

@@
 %%\documentclass[]{spie}  %>>> use for US letter paper
-\documentclass[a4papert,12pt]{spie}  %>>> use this instead for A4 paper
+\documentclass[a4paper,12pt]{spie}  %>>> use this instead for A4 paper
 %%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data Center Profiling } 
-Google data-centers show 50\% less energy consumption than a typical data center.
-Google is able to achieve a trailing twelve-month (TTM) PUE of 1.12 across all data centers, in all seasons, including all sources of overhead.
-These data-centers have varied characteristics over time, which will be discussed in detail in this section.
+\section{Data Center Profiling } \label{sec:datacenter}
+In order to run our experiments, we need to collect information about data centers that will help us make better scheduling decisions.
+We call this process data center profiling.
+The major energy costs in data center are from running the IT resources and from cooling the data centers.
+We create a profile for each of the data center with the information that effects the energy costs.
 \subsection{Geo-Distributed Data Center}
-We have chosen to simulate Google data centers that do the work for us around the world, around the clock at five different locations.
-These consists of two of them in United States namely Berkeley County, South Carolina and Lenoir, North Carolina.
-Other data centers that we have simulated are located one each in Europe, Asia, and South ?America.
+Most cloud service providers including Google, Amazon, Facebook and Apple have data centers distributed across the globe.
+For our experiments, we consider five locations where Google has its data centers\cite{googledatacenter}.
+Table \ref{tab:datacenter} provides the location and labels for each of the data center.
+
+\begin{table}[ht] \label{tab:datacenter}
+\begin{tabular}{lll}
+Location & Label & Associated Trace \\
+Berkeley County, South Carolina & DC1 & A \\
+Lenoir, North Carolina & DC2 & B \\
+Singapore & DC3 & C \\
+Quilicura, Chile & DC4 & D \\
+Dublin, Ireland & DC5 & 
+\end{tabular}
+\end{table}
+
 \subsection{Profile Parameters}
-There are five techniques that Google typically uses to minimize its energy consumption and electricity costs.
-Frequent measuring of Power Usage Effectiveness is the key for determining the power usage for non-computing functions like cooling and power distribution.
-Secondly they manage airflow at data-centers, which allows cooling the hot spots available in data-centers.
-Moreover by setting thermostat to 80$^{\circ}$F it saves significant amount of electricity consumption and costs.
-Google researchers propose maximum usage of ambient temperature through evaporative cooling, cold reservoir, and airside economizer.
-This discards conventional chillers, which are the source of maximum energy consumption at data centers.
-Altogether Google data continuously draw almost 260 million watts, which is about a quarter of the output of a nuclear power plant.
-Google uses renewable energy to power more than 35\% of its operations.
-These are wind, solar energy typically available at all data centers.
+We consider two major parameters that has the most effect on the energy cost.
+They are a) Temperature b) Electricity.
+We collect temperature and electricity prices for these locations from historical archives.
+If time permits, we wish to also consider green energy available at the data center locations.
+We look at each of these parameters and how they effect energy costs in sections \ref{sec:temp}, \ref{sec:elec} and \ref{sec:green}.
 
-Google results show that data centers incur high PUE during summer and low PUE during winters.
-It clearly shows that surrounding temperature plays a critical role in deciding energy consumption at data centers.
-Data centers typically purchase wholesale electricity where the price is determined by day ahead market.
 \begin{figure} %
@@
 
-\subsubsection{Temperature}
-We use temperature as a key parameter to decide the location of job, as temperatures are diverse across data centers and encounters frequent fluctuations in temperature.
-In this case we make our assumption concrete by doing an empirical analysis of historical climate data, which provides substantial evidence.
-We extracted these historical climate data from various data repositories of the National Climate Data Center for all five locations covering the entire one-year period of 2013.
+\subsubsection{Temperature} {sec:temp}
+Data centers located in cold regions have the advantage of using outside air to cool the data center.
+An air-side economizer brings outside air into a building and distributes it to the servers.
+Instead of being re-circulated and cooled, the exhaust air from the servers is simply directed outside.
+This can result in million dollars of saving annually.
+A difference in temperature across data centers mean different calling costs at these data centers.
+We wish to utilize this difference by scheduling jobs in low temperature data centers.
+In this case we make our assumption concrete by doing an empirical analysis of historical climate data.
+We extracted these historical climate data from various data repositories of the National Climate Data Center \cite{cdc} for all five locations covering the entire one-year period of 2013.
 
@@
 
-\subsubsection{Electricity Pricing}
-\subsubsection{Energy Source}
+\subsubsection{Electricity Pricing} {sec:elec}
+Creation of power trading zones have resulted in wholesale power purchase under dynamic pricing schemes.
+Such markets are popular in USA \cite{elecusa}, Germany \cite{elecgermany} and other countries.
+Based on supply and demand bids these there can be high fluctuations in prices \cite{benini2002day}.
+There are many ways to predict there prices \cite{contreras2003arima} \cite{garcia2005garch}.
+If we can predict the electricity prices for the job duration, then we will be able to schedule the job such that the running costs go down.
 
-\subsection{Air Economizer}
+\subsubsection{Energy Source} {sec:green}
+Solar and wind energy are available in most of the modern data centers.
+Similar to electricity price predictions green energy prediction can also be done \cite{goiri2012greenhadoop} \cite{aksanli2012utilizing}.
+Since, this part is in exploratory stage, we will provide more information during the final report.
 
+\subsubsection{Case Study: Google Data Centers} {sec:CSgoogle}
+Google data-centers show 50\% less energy consumption than a typical data center.
+Google is able to achieve a trailing twelve-month (TTM) PUE of 1.12 across all data centers, in all seasons, including all sources of overhead.
+There are five techniques that Google typically uses to minimize its energy consumption and electricity costs.
+Frequent measuring of Power Usage Effectiveness is the key for determining the power usage for non-computing functions like cooling and power distribution.
+Secondly they manage airflow at data-centers, which allows cooling the hot spots available in data-centers.
+Moreover by setting thermostat to 80$^{\circ}$F it saves significant amount of electricity consumption and costs.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Scheduling Algorithm}
-
-Provide basic idea about our strategy
-
-\subsection{Features considered}
-\subsubsection{Energy Cost Model}
-%\subsubsection{Optimization Techiques}
-%\subsubsection{Design}
 

### 2015-04-05 doc/report/midterm/article.tex
tex file update

@@
 %%\documentclass[]{spie}  %>>> use for US letter paper
-\documentclass[a4paper,12pt]{spie}  %>>> use this instead for A4 paper
+\documentclass[a4papert,12pt]{spie}  %>>> use this instead for A4 paper
 %%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
@@
 Today, many of the leading cloud service providers have geographically distributed data centers to serve millions of users around the world.
-The need for a geo-distributed data center arising from the need to provide low-latency and  high-availability of the different services.
-The workloads that run on these data centers are also highly varied.
+The need for a geo-distributed data center arises from the need to provide low-latency and  high-availability of the different services.
+The workloads that run on these data centers are also highly diverse.
 A typical data centers run jobs of different kinds including business-critical workloads---web servers, mail servers, instant messaging services etc, big-data processing, real-time data analytics and HPC jobs.
-As cloud services grow to serve more customers, there is an increasing need for a workload provisioning mechanism which can minimize the cost of operation of the data center while still obeying the SLA and also minimizing the user-perceived latency.
+As cloud services grow to serve more customers, there is an increasing need for a workload provisioning mechanism which can minimize the cost of operation of the data center while minimizing SLA violations and also minimizing the user-perceived latency.
 
 We present GeoSched, a job provisioning framework that is workload aware, energy-aware and cooling-aware.
-GeoSched considers the workload type (batch process, real-time processing, web-service), the energy source of the data center (brown and green energy), electricity pricing of the region, cooling techniques used in the data center (air economizer) and also the history of jobs to predict the load patterns and cluster utilization.
-We wish to evaluate GeoSched on industry workload traces and also on synthetic workloads that can mimic the real-world data center workload.
+GeoSched considers the workload type (batch process, real-time processing, web-service), the energy source of the data center (brown and green energy), electricity pricing of the region and cooling techniques used in the data center (air economizer).
+We GeoSched on industry workload traces and also on synthetic workloads that can mimic the real-world data center workload.
 \end{abstract}
@@
 A good way to study these workloads are by analyzing real-world or synthetic workload traces.
-We look at workload traces in detail in section \ref{traces} and \ref{googletrace}
+We look at cloud workload traces in detail in section \ref{traces} and \ref{googletrace}
 
@@
 \cite{yahootrace}There have also been attempts to generate synthetic workload traces using known real-world workload characteristics \cite{beitch2010rain} \cite{wang2011towards}.
-For the sake of this work, we have decided to use the workload trace from Google.
+For our experiments we have decided to use the workload trace from Google.
 
@@
 \begin {table}[ht]
-\caption {Trace format description} \label{tab:format} 
+\centering
  \scalebox{0.9}{\begin{tabular}{ | p{4cm} |l  |p{11 cm}|  }
@@
     \end{tabular}}
+    \caption {Trace format description} \label{tab:format} 
 \end{table}
@@
 A detailed explanation about the data center profile are given in section \ref{sec:datacenter}.
-To simulate workload execution in these five data centers we have extracted five set of traces each worth five days of incoming workload from the original google trace.
-So, this covers 25 days of traces out of the 29 available days.
+To simulate workload execution in these five data centers we have extracted five set of traces each worth five days of incoming job requests from the original google trace.
+This will cover 25 days of traces out of the 29 available days.
 We label these 5 set of traces from A to E.
-Additionally, in our experiments we wish to only consider those jobs that end without any failures.
-Therefore, we clean the original traces by removing all jobs that is either failed or get killed.
+Additionally, in our experiments we only consider those jobs that end without any failures.
+Therefore, we clean the original traces by removing all jobs that is either fail or get killed.
 
@@
 Generally, these are the kind of jobs that can be provisioned on other data centers as their deadlines are not strict.
-You will find that as the sensitivity increases the running time of these jobs also increase.
+You will find that as the sensitivity increases the running time of these jobs and the resource usage also increase.
 High-sensitive jobs are generally service jobs and run of longer periods of time.
 Figure \ref{fig:pie} shows a pie chart of the job counts in different schedule classes for all the 5 traces.
-You will notice that the job count for class 3 (high latency-sensitivity) jobs are low.
+You will notice that the job count for class 3 (high latency-sensitive) jobs are low.
 This is because these jobs generally run for long period of time, typically months.
@@
 \begin{table}
-\caption {Trace set statistics} \label{tab:tracestat} 
 \adjustbox{max width=1\textwidth}{
@@
 }
+\caption {Trace set statistics} \label{tab:tracestat} 
 \end{table}
 
-Figure \ref{fig:arrivalrate} shows the variation in job arrival for the traces.
+Figure \ref{fig:arrivalrate} shows the variation in job arrival rates for the traces.
 As you can see there is both intra-trace variation as well as inter-trace variation.
@@
 There are opportunities to provision jobs to other data centers if the traffic at a particular data center is high.
-What is also important to notice is that the scheduling decision should be made quickly when the arrival rates of jobs are high, else there are chances for these jobs to miss the SLA.
+What is also important to notice is that the scheduling decision should be made quickly when the arrival rates of jobs are high, else there are chances for that these jobs miss the SLAs.
 Hence, even the scheduler should be scalable during high volume traffic.
@@
 There are a few jobs that have higher run-times.
-As explained earlier, these jobs are primarily internet-service jobs that run for months.
+As explained earlier, these jobs are primarily internet-service jobs that are expected to run for longer duration.
 
@@
 We create a profile for each of the data center with the information that effects the energy costs.
+
+\begin{table}[h] 
+ \centering
+\begin{tabular}{|l|l|l|}
+\hline
+\bf{Location}                        & \bf{Label} & \bf{Associated Trace} \\ \hline
+Berkeley County, South Carolina & DC1   & A                \\ \hline
+Lenoir, North Carolina          & DC2   & B                \\ \hline
+Singapore                       & DC3   & C                \\ \hline
+Quilicura, Chile                & DC4   & D                \\ \hline
+Dublin, Ireland                 & DC5   &   E               \\ \hline
+\end{tabular}
+\caption {Data center location information} \label{tab:dc}
+\end{table}
+
 \subsection{Geo-Distributed Data Center}
 Most cloud service providers including Google, Amazon, Facebook and Apple have data centers distributed across the globe.
-For our experiments, we consider five locations where Google has its data centers\cite{googledatacenter}.
-Table \ref{tab:datacenter} provides the location and labels for each of the data center.
+For our experiments, we consider five locations where Google has its data centers\cite{googletracedata}.
+Table \ref{tab:dc} provides the location and labels for each of the data center.
+We also replicate the computing resource used in the original google trace across all the data centers.
+The configuration of the resources along with the normalzed CPU and memory are given in table \ref{tab:dcres}
 
-\begin{table}[ht] \label{tab:datacenter}
-\begin{tabular}{lll}
-Location & Label & Associated Trace \\
-Berkeley County, South Carolina & DC1 & A \\
-Lenoir, North Carolina & DC2 & B \\
-Singapore & DC3 & C \\
-Quilicura, Chile & DC4 & D \\
-Dublin, Ireland & DC5 & 
+\begin{table}[h] 
+ \centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\bf{Node count}      &       \bf{Architecture}          & \bf{CPU} & \bf{Memory} \\ \hline
+6732 & B & 0.5 & 0.50   \\ \hline
+3863 & B & 0.5 & 0.25   \\ \hline
+1001 & B & 0.5 & 0.75   \\ \hline
+795 & C & 1 & 1   \\ \hline
+126 & A & 0.25 & 0.25   \\ \hline
+52 & B & 0.5 & 0.12  \\ \hline
+5 & B & 0.5 & 0.3  \\ \hline
+5 & B & 0.5 & 0.97  \\ \hline
+3 & C & 1 & 0.50  \\ \hline
+1 & B & 0.5 & 0.06  \\ \hline
 \end{tabular}
+\caption {Data center IT resource configuration} \label{tab:dcres}
 \end{table}
@@
 
-\subsubsection{Temperature} {sec:temp}
+\subsubsection{Temperature} \label{sec:temp}
 Data centers located in cold regions have the advantage of using outside air to cool the data center.
@@
 This can result in million dollars of saving annually.
-A difference in temperature across data centers mean different calling costs at these data centers.
+A difference in temperature across data centers mean different cooling costs at these data centers.
 We wish to utilize this difference by scheduling jobs in low temperature data centers.
 In this case we make our assumption concrete by doing an empirical analysis of historical climate data.
-We extracted these historical climate data from various data repositories of the National Climate Data Center \cite{cdc} for all five locations covering the entire one-year period of 2013.
+We extracted these historical climate data from various data repositories of the National Climate Data Center \cite{climatedata} for all five locations covering the entire one-year period of 2013.
 
@@
 
-\subsubsection{Electricity Pricing} {sec:elec}
+\subsubsection{Electricity Pricing} \label{sec:elec}
 Creation of power trading zones have resulted in wholesale power purchase under dynamic pricing schemes.
 Such markets are popular in USA \cite{elecusa}, Germany \cite{elecgermany} and other countries.
-Based on supply and demand bids these there can be high fluctuations in prices \cite{benini2002day}.
-There are many ways to predict there prices \cite{contreras2003arima} \cite{garcia2005garch}.
+Based on supply and demand bids there can be high fluctuations in prices \cite{benini2002day}.
+There are many ways to predict these prices \cite{contreras2003arima} \cite{garcia2005garch}.
 If we can predict the electricity prices for the job duration, then we will be able to schedule the job such that the running costs go down.
 
-\subsubsection{Energy Source} {sec:green}
+\subsubsection{Energy Source} \label{sec:green}
 Solar and wind energy are available in most of the modern data centers.
@@
 
-\subsubsection{Case Study: Google Data Centers} {sec:CSgoogle}
+\subsubsection{Case Study: Google Data Centers} \label{sec:CSgoogle}
 Google data-centers show 50\% less energy consumption than a typical data center.
@@
 Moreover by setting thermostat to 80$^{\circ}$F it saves significant amount of electricity consumption and costs.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Scheduler simulator} 
+We plan to implement our simulator using CloudSim \cite{calheiros2011cloudsim}.
+CloudSim is an extensible simulation toolkit that enables modeling and simulation of Cloud computing systems and application provisioning environments.
+The CloudSim toolkit supports both system and behavior modeling of Cloud system components such as data centers, virtual machines (VMs) and resource provisioning policies.
+\subsection{Components}
+\begin{list2} %Job Description%
+\item \textbf{Datacenter:} This component will simulate a data center.
+This composed of 1) scheduler that implements a specific provisioning algorithm, 2) data center profile database, 3) a work trace reader and 4)proxy elements of other data center for inter-data center communication\
+\item \textbf{DataCenter Proxy}: All interactions with the data center will be through a proxy. Each datacenter will hold the proxies to other data centers.\
+\item \textbf{Scheduler:} This component will implement the optimization technique and schedule jobs onto data centers.\
+\item \textbf{Datacenter Profile DB}: This components contains the temperature and electricity data. Scheduler will access this to make provisioning decisions.\
+\end{list2}
+
+\begin{figure}[] 
+\centering
+\includegraphics[scale=0.60]{component}
+\caption{Simulation components and their interaction}
+\label{fig:component}
+\end{figure}
+
+\subsubsection{Component Interaction}
+Figure \ref{fig:component} gives the interaction model between the components.
+The datacenter object will start a thread which will run the scheduling algorithm, scheduling jobs from the trace file.
+Each data center had proxies to other data centers if it needs to communicate resource availability or job provisioning to other data centers.
+
+\section{Conclusion}
+We have completed the work trace analysis and data center profiling parts of the project.
+Currently the simulator is being implemented, following which we will work on the energy minimization techniques and the final evaluation of these techniques.
+If time permits, we would also like to consider green energy in our provisioning mechanism.
+Section \ref{sec:mile} shows the current state sand future directions of the project.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
+\section{Scheduling Algorithm}
 
@@
 
-\subsection{Architecture}
-Introduce cloudsim
-\subsubsection{Components}
-\subsubsection{Integration of traces and data center profile}
-\subsubsection{Design}
+\subsection{Features considered}
+\subsubsection{Energy Cost Model}
+%\subsubsection{Optimization Techiques}
+%\subsubsection{Design}
+\fi
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{ Milestones}
-
-    \begin{tabular}{ | p{4cm} |l  |p{11 cm}|  }
+\section{ Milestones} \label{sec:mile}
+\adjustbox{max width=1\textwidth}{
+    \begin{tabular}{ | p{4cm} |l  |p{11 cm}| l |  }
     \hline
-    \bf{Item} & \bf{Deadline} & \bf{Description}   \\ \hline
+    \bf{Item} & \bf{Deadline} & \bf{Description}  &  \bf{Status} \\ \hline
     Study real-world and synthetic workloads & 10\textsuperscript{th} March  & We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.
-\\ \hline
+& Completed  \\ \hline
     Data Center profile & 20\textsuperscript{th} March  & In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-distributed data centers.
-\\ \hline
-    Simulator design and implementation& 12\textsuperscript{th} April  & Implement the simulator infrastructure for multi-datacenter environment.  \\ \hline
-    Feature Extraction and Optimization & 25\textsuperscript{th} April  & Here we identifying all the necessary features that contribute to the energy cost  and build an optimization technique which is adaptive to workload characteristics and data center's resource capacity and electricity costs \\ \hline
-    Evaluation & 7\textsuperscript{th} May  & This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above \\ \hline
+& Completed\\ \hline
+    Simulator design and implementation& 12\textsuperscript{th} April  & Implement the simulator infrastructure for multi-datacenter environment.  & In progress \\ \hline
+    Feature Extraction and Optimization & 25\textsuperscript{th} April  & Here we identifying all the necessary features that contribute to the energy cost  and build an optimization technique which is adaptive to workload characteristics and data center's resource capacity and electricity costs & TBD \\ \hline
+    Evaluation & 7\textsuperscript{th} May  & This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above & TBD \\ \hline
 
     \end{tabular}
+}
 

### 2015-04-13 data_centers/writecsv.py
Added temp csv files

@@ -1,0 +1,114 @@
+import os
+import traceback
+import sys
+import numpy
+import matplotlib.pyplot as plt 
+from pylab import *
+import numpy as np
+import calendar
+
+
+cwd=os.getcwd()
+
+address = cwd+"/"
+IRE=address + "/IRE"
+IRE="/Users/harshitdokania/Desktop/cs525/geosched/data_centers/IRE"
+IRE_file="3982496643522dat.txt"
+NC_file="8845366643370dat.txt"
+NC=address + "/North_Carolina"
+Berk_file="5475956643413dat.txt"
+Berk=address + "/berkeley"
+Chile_file="7361266643506dat.txt"
+Chile=address+"/chile"
+Sing_file="5146466643516dat.txt"
+Sing=address + "/singapore"
+
+
+
+def ConvertCSV(filename, address, output_file):
+ count=0
+ date=0
+ hr=0
+ old_date_0=0
+ old_date_1=1
+ avg=[]
+ sum1=0
+ YEAR=''
+ MONTH=''
+ DAY=''
+ HOUR=''
+ with open(os.path.join(address, filename), 'r') as txt_file:
+     for line in txt_file:
+       myData=line.split() 
+       count=count+1
+       if count==1:
+         continue
+       DATE=myData[2]
+       date_0=DATE[6]
+       date_1=DATE[7]
+       for i in range(0,12):
+         if i<4:
+           YEAR=YEAR+DATE[i]
+         elif i>=4 and i<6:
+           MONTH=MONTH+DATE[i]
+         elif i>=6 and i<8:
+           DAY=DAY + DATE[i]
+         elif i>=8 and i<12:
+           HOUR=HOUR + DATE[i]
+       TEMP=myData[21]
+    #date_0=DATE[6]
+       print DATE
+       formatted=YEAR +',' + MONTH + ',' + DAY + ',' + HOUR + ',' + TEMP 
+       outfile = open(output_file,'a')
+       print >> outfile,formatted
+       outfile.close()
+       YEAR=''
+       MONTH=''
+       DAY=''
+       HOUR=''
+       if count==2:
+         old_date_0=date_0  
+         old_date_1=date_1 
+       if date_0==old_date_0 and date_1==old_date_1 :
[…]

### 2015-04-15 doc/report/Final/usenix.tex
final report double column format

--- 2015-04-05 doc/report/midterm/article.tex
@@
-%  article.tex (Version 3.3, released 19 January 2008)
-%  Article to demonstrate format for SPIE Proceedings
-%  Special instructions are included in this file after the
-%  symbol %>>>>
-%  Numerous commands are commented out, but included to show how
-%  to effect various options, e.g., to print page numbers, etc.
-%  This LaTeX source file is composed for LaTeX2e.
+% TEMPLATE for Usenix papers, specifically to meet requirements of
+%  USENIX '05
+% originally a template for producing IEEE-format articles using LaTeX.
+%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
+% adapted by David Beazley for his excellent SWIG paper in Proceedings,
+%   Tcl 96
+% turned into a smartass generic template by De Clarke, with thanks to
+%   both the above pioneers
+% use at your own risk.  Complaints to /dev/null.
+% make it two column with no page numbering, default is 10 point
 
-%  The following commands have been added in the SPIE class 
-%  file (spie.cls) and will not be understood in other classes:
-%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
-%  The bibliography style file is called spiebib.bst, 
-%  which replaces the standard style unstr.bst.
+% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
+% the .sty file from the LaTeX source template, so that people can
+% more easily include the .sty file into an existing document.  Also
+% changed to more closely follow the style guidelines as represented
+% by the Word sample file.
 
-%%\documentclass[]{spie}  %>>> use for US letter paper
-\documentclass[a4papert,12pt]{spie}  %>>> use this instead for A4 paper
-%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
-%% \addtolength{\voffset}{9mm}   %>>> moves text field down
-%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
-%  The following command loads a graphics package to include images 
-%  in the document. It may be necessary to specify a DVI driver option,
-%  e.g., [dvips], but that may be inappropriate for some LaTeX 
-%  installations. 
+% Note that since 2010, USENIX does not require endnotes. If you want
+% foot of page notes, don't include the endnotes package in the 
+% usepackage command, below.
+
+\documentclass[letterpaper,twocolumn,10pt]{article}
+\usepackage{usenix,epsfig,endnotes}
 \usepackage[]{graphicx}
@@
 \usepackage{adjustbox,lipsum}
-\graphicspath{ {/Users/anirudhnair/code/GeoSched/src/GoogleTrace/plots} }
 \usepackage{subfig}
-\usepackage[margin=0.5in]{geometry}
 
-\title{GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers} 
-%\author{Anirudh Jayakumar and Harshit Dokania \\
-%University of Illinois at Urbana-Champaign 
-%\{ajayaku2,hdokani2\}@illinois.edu 
-%}
+\title{\Large \bf GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers} 
+\author{
+{\rm Anirudh Jayakumar}
+\hspace{20 mm}
+{\rm Harshit Dokania}\\
+University of Illinois at Urbana-Champaign
+}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-%>>>> uncomment following for page numbers
-% \pagestyle{plain}    
-%>>>> uncomment following to start page numbering at 301 
-%\setcounter{page}{301}
-
   \begin{document} 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\begin{abstract}
+\subsection*{ABSTRACT}
 Today, many of the leading cloud service providers have geographically distributed data centers to serve millions of users around the world.
@@
 We GeoSched on industry workload traces and also on synthetic workloads that can mimic the real-world data center workload.
-\end{abstract}
+
+\section{INTRODUCTION}
+\label{sec:intro}
+
+\textbf{Problem Statement:} Minimize the total cost of operations of a geo-distributed data center running cloud workloads by minimizing the energy costs subject to business constraints like the service level agreement(SLA) and other technical constraints like data center capacity limits.
+The energy costs will depend on the power usage of the IT infrastructure and also the energy required by the cooling infrastructure to maintain the desired temperature inside the data center.
+Additionally, the solution should also consider the different energy source(green and brown) available at the data center and the different workload types that are typical to a cloud data center.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:intro}  % \label{} allows reference to this section
+\section{RELATED WORK}
+\label{sec:related} 
 Cost minimization has been extensively studied both in the context of a single data center and a geo-distributed data center.
@@
 One such approach is proposed by Rodero et al.
-\cite{rodero2012energy} where the author  propose to use applications۪s profiling (CPU, memory, bandwidth) to  design application-centric energy-aware strategy for VM allocation during VM migration.
+\cite{rodero2012energy} where the author  propose to use applicationss profiling (CPU, memory, bandwidth) to  design application-centric energy-aware strategy for VM allocation during VM migration.
 
@@
 \cite{goiri2011greenslot} propose a system GreenSlot that maximizes the use of green energy available at data centers while ensuring all job deadlines are met.
-GreenSlot predicts the solar energy available in data centers in near future and schedules the workload to maximize green energy consumption and meeting jobs۪s deadlines.
+GreenSlot predicts the solar energy available in data centers in near future and schedules the workload to maximize green energy consumption and meeting jobss deadlines.
 Green- Slot can increase green energy consumption by up to 117\% and decrease energy cost by up to 39\%, compared to a conventional scheduler.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Problem Statement}
-
-Minimize the total cost of operations of a geo-distributed data center running cloud workloads by minimizing the energy costs subject to business constraints like the service level agreement(SLA) and other technical constraints like data center capacity limits.
-The energy costs will depend on the power usage of the IT infrastructure and also the energy required by the cooling infrastructure to maintain the desired temperature inside the data center.
-Additionally, the solution should also consider the different energy source(green and brown) available at the data center and the different workload types that are typical to a cloud data center.
+\fi
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\fi
 
-\section{Cloud Workload } 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{CLOUD WORKLOADS } 
 With the increase in the adoption of cloud computing by businesses around the world, more and more workloads are moving out of traditional data centers into the cloud.
@@
 
-\begin {table}[ht]
+\begin {table}[]
 \centering
- \scalebox{0.9}{\begin{tabular}{ | p{4cm} |l  |p{11 cm}|  }
+ \scalebox{0.7}{\begin{tabular}{ | p{3cm} |l  |p{6 cm}|  }
     \hline
     \bf{Field Name} & \bf{Type} & \bf{Description}   \\ \hline
-     JobID & INTEGER  & Unique job identifier for each job\\ \hline
-     Arrival Time & INTEGER & Arrival time in microseconds \\ \hline
-     Scheduled Time & INTEGER  & Arrival time in microseconds\\ \hline
-     Finish Time & INTEGER  & Arrival time in microseconds\\ \hline
-     Schedule class & INTEGER  & Latency class of the job. \\ \hline
-     Tasks & INTEGER  & Total number of tasks in the job. \\ \hline
-     Total CPU& INTEGER  & Aggregate sum of CPU used by all the tasks in the job\\ \hline
-     Total Memory& INTEGER   & Aggregate sum of memory used by all the tasks in the job \\ \hline
-     User& INTEGER  & Unique user identifier \\ \hline
+     JobID & INT  & Unique job identifier for each job\\ \hline
+     Arrival Time & INT & Arrival time in microseconds \\ \hline
+     Scheduled Time & INT  & Arrival time in microseconds\\ \hline
+     Finish Time & INT  & Arrival time in microseconds\\ \hline
+     Schedule class & INT  & Latency class of the job. \\ \hline
+     Tasks & INT  & Total number of tasks in the job. \\ \hline
+     Total CPU& INT  & Aggregate sum of CPU used by all the tasks in the job\\ \hline
+     Total Memory& INT   & Aggregate sum of memory used by all the tasks in the job \\ \hline
+     User& INT  & Unique user identifier \\ \hline
      SLA Param1&  FLOAT & SLA parameter placeholder\\ \hline
@@
 
-\begin{figure}[ht] 
+\begin{figure}[] 
 \centering
-\includegraphics[scale=0.40]{arrivalrate.png}
+\includegraphics[scale=0.25]{arrivalrate.png}
 \caption{Arrival rate of jobs }
@@
 
-\begin{figure}[]
+\begin{figure*}[]
 \centering
@@
  \label{fig:pie}
-\end{figure}
+\end{figure*}
 
-\begin{table}
+\begin{table*}
 \adjustbox{max width=1\textwidth}{
@@
 \caption {Trace set statistics} \label{tab:tracestat} 
-\end{table}
+\end{table*}
 
@@
 \centering
-\includegraphics[scale=0.40]{runtime}
+\includegraphics[scale=0.25]{runtime}
 \caption{CDF of the job duration}
@@
     \centering
-    \subfloat[CPU resource ]{{\includegraphics[width=9cm]{cpu} }}%
-    \subfloat[RAM resource]{{\includegraphics[width=9cm]{mem} }}%
+    \subfloat[CPU resource ]{{\includegraphics[width=8cm]{cpu} }}%
+    \newline
+
+    \subfloat[RAM resource]{{\includegraphics[width=8cm]{mem} }}%
     \caption{CDF of resource requested by jobs}%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data Center Profiling } \label{sec:datacenter}
+\section{DATA CENTER PROFILING } \label{sec:datacenter}
 In order to run our experiments, we need to collect information about data centers that will help us make better scheduling decisions.
@@
 
-\begin{table}[h] 
+\begin{table}[] 
  \centering
-\begin{tabular}{|l|l|l|}
+\begin{tabular}{|p{3cm}|l|p{2.5cm}|}
 \hline
@@
 
-\begin{table}[h] 
+\begin{table}[] 
  \centering
-\begin{tabular}{|l|l|l|l|}
+\scalebox{0.8}{\begin{tabular}{|l|l|l|l|}
 \hline
@@
 1 & B & 0.5 & 0.06  \\ \hline
-\end{tabular}
+\end{tabular}}
 \caption {Data center IT resource configuration} \label{tab:dcres}
@@
 
-\begin{figure} %
+\begin{figure*} %
     \centering
-    \subfloat[Daily Average Temperature ]{{\includegraphics[width=9cm]{AverageTemp} }}%
-    \subfloat[Daily Maximum Variation]{{\includegraphics[width=9cm]{Variation} }}%
+    \subfloat[Daily Average Temperature ]{{\includegraphics[width=8cm]{AverageTemp} }}%
+    \subfloat[Daily Maximum Variation]{{\includegraphics[width=8cm]{Variation} }}%
     \caption{Temperature variation for the year 2013}%
     \label{fig:temp}%
-\end{figure}
+\end{figure*}
 
@@
 
-\section{Scheduler simulator} 
+\section{SCHEDULER SIMULATOR} 
 We plan to implement our simulator using CloudSim \cite{calheiros2011cloudsim}.
@@
 \centering
-\includegraphics[scale=0.60]{component}
+\includegraphics[scale=0.40]{component}
 \caption{Simulation components and their interaction}
@@
 
-\section{Conclusion}
+\section{CONCLUSION}
 We have completed the work trace analysis and data center profiling parts of the project.
@@
 If time permits, we would also like to consider green energy in our provisioning mechanism.
-Section \ref{sec:mile} shows the current state sand future directions of the project.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \iffalse
-\section{Scheduling Algorithm}
+\section{SCHEDULING ALGORITHM}
 
@@
 
-\section{ Milestones} \label{sec:mile}
-\adjustbox{max width=1\textwidth}{
-    \begin{tabular}{ | p{4cm} |l  |p{11 cm}| l |  }
-    \hline
-    \bf{Item} & \bf{Deadline} & \bf{Description}  &  \bf{Status} \\ \hline
-    Study real-world and synthetic workloads & 10\textsuperscript{th} March  & We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.
-& Completed  \\ \hline
-    Data Center profile & 20\textsuperscript{th} March  & In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-distributed data centers.
-& Completed\\ \hline
-    Simulator design and implementation& 12\textsuperscript{th} April  & Implement the simulator infrastructure for multi-datacenter environment.  & In progress \\ \hline
-    Feature Extraction and Optimization & 25\textsuperscript{th} April  & Here we identifying all the necessary features that contribute to the energy cost  and build an optimization technique which is adaptive to workload characteristics and data center's resource capacity and electricity costs & TBD \\ \hline
-    Evaluation & 7\textsuperscript{th} May  & This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above & TBD \\ \hline
-
-    \end{tabular}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliography{report}   %>>>> bibliography data in report.bib
-\bibliographystyle{spiebib}   %>>>> makes bibtex use spiebib.bst
+{\footnotesize \bibliographystyle{acm}
+\bibliography{sample}}
 

### 2015-05-08 src/GeoSim/GoogleTrace.cpp
bug fixes
Change-Id: Id008ef7545215c038e239488a5dd6e43a9d5f2e2

@@ -82,8 +82,9 @@ INT64_ TraceItem::ArrivalTime()
 	return nArrivalTime;
 }
 Job* TraceItem::createJob(){
-
+    //cout << "=====creating jobs " << nJobID << "  " << nTasks  << "======"<< endl; 
 	Job* nJob= new Job(nJobID, RunningTime(), nSchedClass, nTasks,fTotalCPU, fTotalMem );
+    cout << GenerateCSV() << endl;
 	return nJob;
 
 }

### 2015-05-08 src/GeoSim/Job.h

@@ -29,8 +29,6 @@ public:
 	void setNode(Node* node);
 	Job* getJob();
 	void setJob(Job* job);
-    map<int, int> getPossibleNodes();
-    int setPossibleNodes(map<int,int> pN);
 
 	//add required methods
 private:
@@ -39,26 +37,23 @@ private:
 	int cpu;
 	Node *pNode;
 	Job *pJob;
-	map<int,int> possibleNode;
 };
 
 class Job {
 public:
+    Job();
 	Job(INT64_, INT64_, int,  int , double , double);
 	virtual ~Job();
 	INT64_ getJobID();
 	int sClass();
-	int setRetireTime();
-	int setCurrTime(INT64_ arrival);
-	INT64_ getRetireTime();
 	INT64_ getCurrTime();
 	int  IncCurrTime(INT64_ timestamp);
-	std::vector<Task*>& getTasks();
-	// add required method
+	std::vector<Task*> getTasks();
+	INT64_ GetTotalRunTime();
+    // add required method
 private:
 	std::vector<Task*> vTasks;
 	INT64_ nTotalRunTime;
-	INT64_ nRetireTime;
 	INT64_ nCurrRunTime;
 	INT64_ nJobID;
 	int    nSchedClass;

### 2015-05-08 src/GeoSim/common.cpp

@@ -14,19 +14,22 @@ using namespace std;
 
 Barrier::Barrier(std::size_t count)
 {
+    thread_count = 0;
 	_count = _org_val = count;
 }
 
-void Barrier::Wait()
+void Barrier::Wait(int local_count)
 {
+    
 	std::unique_lock<std::mutex> lock{_mutex};
 	if (--_count == 0) {
-		cout << "barrier:" << _count << endl; 
+		//cout << "barrier:" << _count << endl; 
 		_count = _org_val;
+        thread_count = local_count;
 		_cv.notify_all();
 	} else {
-		cout << "barrier:" << _count << endl; 
-		_cv.wait(lock, [this] { return _count == _org_val; });
+		//cout << "barrier:" << _count << endl; 
+		_cv.wait(lock, [this,local_count] { return thread_count == local_count; });
 	}
 }
 

### 2015-05-08 src/GeoSim/common.h

@@ -35,9 +35,10 @@ private:
     std::condition_variable _cv;
     std::size_t _count;
     std::size_t _org_val;
+    int  thread_count;
 public:
     explicit Barrier(std::size_t count);
-    void Wait();
+    void Wait(int);
 };
 
 

### 2015-05-08 src/GeoSim/DataCenter.cpp
bug fixes
Change-Id: Id008ef7545215c038e239488a5dd6e43a9d5f2e2

@@ -1,4 +1,5 @@
 /*
+
  * DataCenter.cpp
  *
  *  Created on: May 2, 2015
@@ -79,7 +80,7 @@ void DataCenter::AddJobsToWaitingList(Job *pJob) {
 	m_waitMutex.lock();
 	m_vWaitingJobs.push_back(pJob);
 	m_waitMutex.unlock();
-	cout << "Added to wait list\n";
+	//cout << "Added to wait list\n";
 
 }
 
@@ -110,7 +111,7 @@ void DataCenter::StartSimulation() {
 
 	cout << "Simulation Started" << endl;
 	INT64_ arrivalTime = TIMEINC;
-
+    int cycle_count = 1;
 	// loop till the time there are jobs left in the trace
 	// or till the time all jobs complete running.
 	while (!m_workLoad.FileEnd() || m_vRunningJobs.size()) {
@@ -119,7 +120,7 @@ void DataCenter::StartSimulation() {
         
         vector<TraceItem*> currLoad = m_workLoad.GetNextSet(arrivalTime);
 		if (currLoad.size()) {
-			cout << "DC " << nDCid << ": Job set size: " <<  currLoad.size() << endl;
+			//cout << "DC " << nDCid << ": Job set size: " <<  currLoad.size() << endl;
 			for (auto it = currLoad.begin(); it != currLoad.end(); it++) {
 				//
 				Job* pJob = (*it)->createJob();
@@ -127,15 +128,15 @@ void DataCenter::StartSimulation() {
 				/* high sensitive jobs put in my own queue */
 				if (pJob->sClass() == 2 || pJob->sClass() == 3) {
 					//	cout<<"job = "<<pJob->sClass()<<endl;
-					cout << "DC " << nDCid << ": JobID: " <<  pJob->getJobID() << " send to local queue due to high sensitivity" << endl;
+					//cout << "DC " << nDCid << ": JobID: " <<  pJob->getJobID() << " send to local queue due to high sensitivity" << endl;
 					AddJobsToWaitingList(pJob); // schedule during schedule phase
 				} else {
 
 					auto vecDCs = GetDCSchedulable(pJob);
-					cout << "DC " << nDCid << ": JobID: " <<  pJob->getJobID() << " Schedulable DCs";
-					for (auto iter = vecDCs.begin(); iter != vecDCs.end(); ++iter)
[…]

### 2015-05-08 src/GeoSim/Job.cpp

@@ -8,8 +8,11 @@
 #include "Job.h"
 #include "Node.h"
 #include "common.h"
+using namespace std;
 
-
+Job::Job()
+{
+}
 
 Job::Job(INT64_ id, INT64_ runtime, int sClass, int tasks, double cpu, double mem) {
 	// TODO Auto-generated constructor stub
@@ -20,7 +23,7 @@ Job::Job(INT64_ id, INT64_ runtime, int sClass, int tasks, double cpu, double me
    double myCpu = cpu/tasks;
    double myMem = mem/tasks;
    for(int i=0; i<tasks ; i++){
-	   Task *t= new Task;
+	   Task *t= new Task();
 	   t->setMem(myMem);
 	   t->setCpu(myCpu);
 	   t->setJob(this);
@@ -28,16 +31,21 @@ Job::Job(INT64_ id, INT64_ runtime, int sClass, int tasks, double cpu, double me
    }
 }
 
-std::vector<Task*>& Job:: getTasks(){
- return vTasks;
+vector<Task*> Job:: getTasks(){
+ vector<Task*> tasks = vTasks;
+ return tasks;
 }
 
 Job::~Job() {
-	for(auto iter = vTasks.begin(); iter != vTasks.end(); ++iter)
+	for(auto iter = vTasks.begin(); iter != vTasks.end();)
 	{
 		Task *pTask = *iter;
+        iter++;
 		delete pTask;
+
 	}
+    vTasks.clear();
+    
 }
 INT64_ Job:: getJobID(){
 	return nJobID;
@@ -47,20 +55,11 @@ int Job::sClass(){
 	return nSchedClass;
 }
 
-
-int Job:: setRetireTime(){
-	nRetireTime= nCurrRunTime + nTotalRunTime;
-	return SUCCESS;
-}
-int Job:: setCurrTime(INT64_ arrival){
-		nCurrRunTime = arrival;
-		setRetireTime();
-		return SUCCESS;
+INT64_ Job::GetTotalRunTime()
+{
+    return nTotalRunTime;
 }
 
-INT64_ Job:: getRetireTime(){
-	return nRetireTime;
-}
 
 INT64_ Job:: getCurrTime(){
 	return nCurrRunTime;
@@ -114,13 +113,6 @@ void Task::setNode(Node* node){
  Job* Task::getJob(){
 	return pJob;
 }
- map<int, int> Task::getPossibleNodes(){
-	 return possibleNode;
[…]

### 2015-05-08 src/GeoSim/DataCenter.cpp
Timezone, Thread exit done, Log still clutterred, util pending

@@ -15,19 +15,25 @@
 #include <fstream>
 #include <iostream>
 #include "ConfigAccessor.h"
+#include <stdio.h>
 
 using namespace std;
 
 #define TIMEINC 				5000000   // 5 secs in us
 #define RESOURCE_SYNC_TIME		900000000 // 900 secs in us
 
-const string machineConfig("../../datacenters/machine_config.csv");
+const string machineConfig("/Users/harshitdokania/Desktop/cs525/geosched/datacenters/machine_config.csv");
 
-DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier) {
+DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string n, int gmtDiff) {
 	// TODO Auto-generated constructor stub
 	nDCid = id;
 	m_sWorkloadTrace = workloadPath;
 	m_pBarrier = pBarrier;
+    name = n;
+    m_GMT = gmtDiff;
+    
+    string s= "Started creating " + name + " DataCenter at " +getLocalTime() + "\n";
+    L.print(s);
 
 }
 
@@ -46,6 +52,7 @@ int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
 	int memAvail;
 	ifstream fileStream(machineConfig);
 	int nodeID = 0;
+   
 
 	NodeMap mapNodes;
 	while (!fileStream.eof()) {
@@ -68,16 +75,21 @@ int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
 		}
 	}
 	m_mapDCtoResource[nDCid] = mapNodes;
+    string s= "Completed creating "+ name +" DataCenter at "+ getLocalTime()+"\n";
+    L.print(s);
+    
+    
 
-	//cout<<"Started tracing"<<endl;
+	
 
-	m_workLoad.Initialize(m_sWorkloadTrace);
+	m_workLoad.Initialize(m_sWorkloadTrace, name, m_GMT);
 
 	return SUCCESS;
 }
 
 void DataCenter::AddJobsToWaitingList(Job *pJob) {
-	m_waitMutex.lock();
+   
+    m_waitMutex.lock();
 	m_vWaitingJobs.push_back(pJob);
 	m_waitMutex.unlock();
 	//cout << "Added to wait list\n";
@@ -96,7 +108,20 @@ list<Job*> DataCenter::GetWaitingJobs() {
 
 }
 
[…]

### 2015-05-08 src/GeoSim/DataCenterProxy.cpp

@@ -29,10 +29,23 @@ int DataCenterProxy:: Initialize(DataCenter *dc)
 
 int DataCenterProxy::SubmitJob(Job* pJob)
 {
-	cout<<"Submitting job\n";
+	
 	m_dc->AddJobsToWaitingList(pJob);
 	return SUCCESS;
 }
+int DataCenterProxy::GetName()
+{
+    
+    m_dc->GetName();
+    return SUCCESS;
+}
+
+void DataCenterProxy::InformLeaving(){
+    
+    m_dc->decrementBarrier();
+}
+
+
 
 std::unordered_map<int,Node*> DataCenterProxy:: GetResourceData()
 {

### 2015-05-08 src/GeoSim/GoogleTrace.cpp

@@ -88,6 +88,32 @@ Job* TraceItem::createJob(){
 	return nJob;
 
 }
+string  localtime(int i){
+    
+    string Time;
+    time_t m_Time;
+    struct tm * ptm;
+    
+    time ( &m_Time );
+    ptm = gmtime(&m_Time);
+    if(ptm->tm_hour+i<=0){
+        Time=to_string(ptm->tm_hour+i+12);
+    }
+    else
+        Time=to_string((ptm->tm_hour+i)%24);
+    Time+=":";
+    if((ptm->tm_min)<9) {
+        Time+="0" ;
+    }
+    Time+=to_string((ptm->tm_min));
+    Time+=":";
+    if((ptm->tm_sec)<9){
+        Time+="0" ;
+    }
+    Time+=to_string((ptm->tm_sec));
+    return Time;
+}
+
 
 std::string TraceItem::GenerateCSV()
 {
@@ -109,7 +135,7 @@ std::string TraceItem::GenerateCSV()
 }
 
 
-int GoogleTrace::Initialize(const std::string &filePath)
+int GoogleTrace::Initialize(const std::string &filePath, string name, int GMT)
 {
 
 	ifstream traceStream(filePath);
@@ -117,19 +143,19 @@ int GoogleTrace::Initialize(const std::string &filePath)
 	while(!traceStream.eof())
 	{
 		traceStream >> sLine;
-    //    cout<<sLine<<"\n";
+    
 		TraceItem job_;
 		job_.Init(sLine);
 		vTraceItems.push_back(job_);
-		//cout<<"Tracing"<<endl;
 	}
 	currIter = vTraceItems.begin();
-
+    
 	traceStream.close();
+    cout<<name<<" Completed Workload tracing at "<<localtime(GMT)<<endl;
 	return SUCCESS;
 }
 
-vector<TraceItem *> GoogleTrace::GetNextSet(INT64_ us)
+vector<TraceItem *> GoogleTrace::GetNextSet(INT64_ us, string name, int GMT)
 {
 	vector<TraceItem *> ret;
 	if(currIter == vTraceItems.end())
@@ -144,6 +170,7 @@ vector<TraceItem *> GoogleTrace::GetNextSet(INT64_ us)
 			break;
 		nextJob = &(*currIter);
 	}
+    
 	return ret;
 }
 

### 2015-05-08 src/GeoSim/common.cpp

@@ -33,6 +33,15 @@ void Barrier::Wait(int local_count)
 	}
 }
 
+void Barrier:: decrement(){
+    --_org_val;
+}
+
+void Log:: print(string s){
+    m_LogMutex.lock();
+    cout<<s;
+    m_LogMutex.unlock();
+}
 
 vector<string> &Tokenize(const string &s, char delim, vector<string> &elems)
 {
@@ -54,3 +63,4 @@ vector<int> &Tokenize(const string &s, char delim, vector<int> &elems)
 	}
 	return elems;
 }
+

### 2015-05-08 src/GeoSim/main.cpp

@@ -12,45 +12,52 @@
 #include "DataCenterProxy.h"
 #include "ConfigAccessor.h"
 using namespace std;
-const string workload_chile("../../workloads/google/5_1.csv");
-const string workload_finland("../../workloads/google/5_2.csv");
-const string workload_singapore("../../workloads/google/5_3.csv");
-const string workload_oregon("../../workloads/google/5_4.csv");
-const string workload_iowa("../../workloads/google/5_5.csv");
+const string workload_chile("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
+const string workload_finland("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
+const string workload_singapore("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
+const string workload_oregon("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
+const string workload_iowa("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
 const string config_xml("config.xml");
+
+
+
+
 
 
 int main()
 {
-	//Barrier(5);
-	cout << "Running" << endl;
+	
+	cout << "GeoSim Simulation" << endl;
 	
 	Barrier oBarrier(DC_COUNT);
 
-	DataCenter chile(CHILE, workload_chile,&oBarrier);
-	DataCenter finland(FINLAND, workload_finland,&oBarrier);
-	DataCenter singapore(SINGAPORE, workload_singapore,&oBarrier);
-	DataCenter oregon(OREGON, workload_oregon,&oBarrier);
-	DataCenter iowa(IOWA, workload_iowa,&oBarrier);
+	DataCenter chile(CHILE, workload_chile,&oBarrier, "Chile", -3);
+	DataCenter finland(FINLAND, workload_finland,&oBarrier, "Finland",3);
+	DataCenter singapore(SINGAPORE, workload_singapore,&oBarrier, "Singapore", -4);
+	DataCenter oregon(OREGON, workload_oregon,&oBarrier, "Oregon", -7);
+	DataCenter iowa(IOWA, workload_iowa,&oBarrier,"Iowa", -5);
 	
+    
+    
 	DataCenterProxy *dcProxy= new DataCenterProxy[DC_COUNT];
[…]

### 2015-05-08 src/GeoSim/DataCenter.h
utilization pending and log files still cluttered

@@ -33,7 +33,7 @@ typedef std::unordered_map<int,Node*>                NodeMap;
 class Job;
 class DataCenter {
 public:
-	DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier);
+	DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string name, int gmtDiff);
 	virtual ~DataCenter();
 	int 			Initialize(DataCenterProxy * dataCenterProxies, ConfigAccessor *pAccessor);
 	void 			AddJobsToWaitingList(Job *pJob);
@@ -42,9 +42,14 @@ public:
 	NodeMap 		GetResourceData();
 	void 			UpdateResourceData();
 	void 			set_dataCenterProxies(DataCenterProxy *proxy);
+    void            DecreaseBarrier();
 	GoogleTrace* 	getWorkLoad();
 	void 			Join();
 	void 			Simulation();
+    string          getLocalTime();
+    string          GetName();
+    Barrier*        getBarrier();
+    void            decrementBarrier();
 
 
 private:
@@ -61,7 +66,7 @@ private:
 	std::string			m_sWorkloadTrace;
 	Barrier				*m_pBarrier;
 	ConfigAccessor      *m_pAccessor;
-private:
+    int                 m_GMT;
 	bool				CheckDCFit(Job *,NodeMap &);
 	void 				StartSimulation(); //main scheduling logic goes in here
 	std::vector<int>    GetDCSchedulable(Job *);
@@ -70,6 +75,10 @@ private:
 	int 				ScheduleJob(Job *);
 	void 				RemoveJobFromWaitingQueue(Job*);
 	std::list<Job*> 	GetWaitingJobs();
+    string              localtime(int gmtDiff);
+    string              name;
+     Log                 L;
+   
 };
 
 #endif /* DATACENTER_H_ */

### 2015-05-08 src/GeoSim/DataCenterProxy.h

@@ -23,6 +23,8 @@ public:
 	int Initialize(DataCenter *dc);
 	int SubmitJob(Job* pJob);
 	std::unordered_map<int,Node*> GetResourceData();
+    void InformLeaving();
+    int GetName();
 private:
 	DataCenter *m_dc;
 };

### 2015-05-08 src/GeoSim/common.h

@@ -38,8 +38,23 @@ private:
     int  thread_count;
 public:
     explicit Barrier(std::size_t count);
+    void decrement();
     void Wait(int);
 };
+
+
+class Log{
+    
+private:
+    std::mutex  m_LogMutex;
+public:
+   void print(string s);
+};
+
+
+
+
+
 
 
 

### 2015-05-09 src/GeoSim/DataCenter.cpp
Utilization still 0

@@ -24,18 +24,20 @@ using namespace std;
 
 const string machineConfig("/Users/harshitdokania/Desktop/cs525/geosched/datacenters/machine_config.csv");
 
-DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string n, int gmtDiff) {
+DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string n, int gmtDiff, string tracepath) {
 	// TODO Auto-generated constructor stub
 	nDCid = id;
 	m_sWorkloadTrace = workloadPath;
 	m_pBarrier = pBarrier;
     name = n;
     m_GMT = gmtDiff;
-    
-    string s= "Started creating " + name + " DataCenter at " +getLocalTime() + "\n";
-    L.print(s);
+    m_ExecutionTraces = tracepath;
 
-}
+    
+    string s= "Started creating " + name + " DataCenter at " +getLocalTime();
+   
+    Logfile(s);
+   }
 
 DataCenter::~DataCenter() {
 	// TODO Auto-generated destructor stub
@@ -75,14 +77,13 @@ int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
 		}
 	}
 	m_mapDCtoResource[nDCid] = mapNodes;
-    string s= "Completed creating "+ name +" DataCenter at "+ getLocalTime()+"\n";
-    L.print(s);
-    
-    
+    string s= "Completed creating "+ name +" DataCenter at "+ getLocalTime();
+    s= "Debug utilize while creating ";
+    Logfile(s);
+    PrintUtilization();
 
-	
-
-	m_workLoad.Initialize(m_sWorkloadTrace, name, m_GMT);
+    Logfile(s);
+    m_workLoad.Initialize(m_sWorkloadTrace, name, m_GMT, m_ExecutionTraces);
 
 	return SUCCESS;
 }
@@ -97,6 +98,43 @@ void DataCenter::AddJobsToWaitingList(Job *pJob) {
 }
 
 void DataCenter::PrintUtilization() {
+    NodeMap &nodeMap = m_mapDCtoResource[nDCid];
+    INT64_ totalCpu(0);
+    INT64_ freeCpu(0);
+    INT64_ useCpu(0);
+    INT64_ totalMem(0);
+    INT64_ freeMem(0);
+    INT64_ useMem(0);
+    double pCpu;
+    double pMem;
+
+    
[…]

### 2015-05-09 src/GeoSim/DataCenter.h

@@ -33,7 +33,7 @@ typedef std::unordered_map<int,Node*>                NodeMap;
 class Job;
 class DataCenter {
 public:
-	DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string name, int gmtDiff);
+	DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string name, int gmtDiff, string traces);
 	virtual ~DataCenter();
 	int 			Initialize(DataCenterProxy * dataCenterProxies, ConfigAccessor *pAccessor);
 	void 			AddJobsToWaitingList(Job *pJob);
@@ -77,7 +77,10 @@ private:
 	std::list<Job*> 	GetWaitingJobs();
     string              localtime(int gmtDiff);
     string              name;
-     Log                 L;
+    Log                 L;
+    string              m_ExecutionTraces;
+    int Logfile(string msg);
+   
    
 };
 

### 2015-05-09 src/GeoSim/DataCenterProxy.cpp

@@ -33,11 +33,11 @@ int DataCenterProxy::SubmitJob(Job* pJob)
 	m_dc->AddJobsToWaitingList(pJob);
 	return SUCCESS;
 }
-int DataCenterProxy::GetName()
+string DataCenterProxy::GetName()
 {
     
-    m_dc->GetName();
-    return SUCCESS;
+    return m_dc->GetName();
+    
 }
 
 void DataCenterProxy::InformLeaving(){

### 2015-05-09 src/GeoSim/DataCenterProxy.h

@@ -24,7 +24,7 @@ public:
 	int SubmitJob(Job* pJob);
 	std::unordered_map<int,Node*> GetResourceData();
     void InformLeaving();
-    int GetName();
+    std::string GetName();
 private:
 	DataCenter *m_dc;
 };

### 2015-05-09 src/GeoSim/GoogleTrace.cpp

@@ -84,7 +84,7 @@ INT64_ TraceItem::ArrivalTime()
 Job* TraceItem::createJob(){
     //cout << "=====creating jobs " << nJobID << "  " << nTasks  << "======"<< endl; 
 	Job* nJob= new Job(nJobID, RunningTime(), nSchedClass, nTasks,fTotalCPU, fTotalMem );
-    cout << GenerateCSV() << endl;
+   // cout << GenerateCSV() << endl;
 	return nJob;
 
 }
@@ -135,7 +135,7 @@ std::string TraceItem::GenerateCSV()
 }
 
 
-int GoogleTrace::Initialize(const std::string &filePath, string name, int GMT)
+int GoogleTrace::Initialize(const std::string &filePath, string name, int GMT, string path)
 {
 
 	ifstream traceStream(filePath);
@@ -151,9 +151,25 @@ int GoogleTrace::Initialize(const std::string &filePath, string name, int GMT)
 	currIter = vTraceItems.begin();
     
 	traceStream.close();
-    cout<<name<<" Completed Workload tracing at "<<localtime(GMT)<<endl;
+    string s= name+ " Completed Workload tracing at "+ localtime(GMT);
+    Logfile(s,path);
 	return SUCCESS;
 }
+
+int GoogleTrace:: Logfile(string msg, string path)
+{
+    
+    std::fstream execTraces;
+    execTraces.open (path, std::fstream::in | std::fstream::out | std::fstream::app);
+    
+    execTraces << msg<<endl;
+    
+    execTraces.close();
+    
+    
+    return SUCCESS;
+}
+
 
 vector<TraceItem *> GoogleTrace::GetNextSet(INT64_ us, string name, int GMT)
 {

### 2015-05-09 src/GeoSim/GoogleTrace.h

@@ -55,11 +55,13 @@ private:
 	std::vector<TraceItem>::iterator currIter;
 public:
 
-	int Initialize(const std::string &filePath, string name, int GMT);
+	int Initialize(const std::string &filePath, string name, int GMT, string path);
 	std::vector<TraceItem*> GetNextSet(INT64_ us, string name, int GMT);
 	bool  FileEnd();
 	int WriteToFile(const std::string &filePath);
-};
+    int Logfile(string msg, string path);
+    };
+
 
 
 #endif /* GOOGLETRACE_H_ */

### 2015-05-09 src/GeoSim/Job.cpp

@@ -51,6 +51,17 @@ INT64_ Job:: getJobID(){
 	return nJobID;
 }
 
+int Job:: setNodeID(int n){
+    nodeID = n;
+    return SUCCESS;
+    
+}
+
+int Job:: getNodeID(){
+    return nodeID;
+    
+}
+
 int Job::sClass(){
 	return nSchedClass;
 }

### 2015-05-09 src/GeoSim/Job.h

@@ -45,6 +45,8 @@ public:
 	Job(INT64_, INT64_, int,  int , double , double);
 	virtual ~Job();
 	INT64_ getJobID();
+    int setNodeID(int n);
+    int getNodeID();
 	int sClass();
 	INT64_ getCurrTime();
 	int  IncCurrTime(INT64_ timestamp);
@@ -56,6 +58,7 @@ private:
 	INT64_ nTotalRunTime;
 	INT64_ nCurrRunTime;
 	INT64_ nJobID;
+    int nodeID;
 	int    nSchedClass;
 };
 

### 2015-05-09 src/GeoSim/Node.cpp

@@ -18,9 +18,15 @@ int nTotalCPU;
 	int nNodeID;
 	std::vector<Task*> vTasks;
 
-	int Node:: getFreeCPU(){
+int Node:: getFreeCPU(){
 		return nFreeCPU;
-	}
+}
+int Node:: getTotalCPU(){
+    return nTotalCPU;
+}
+int Node:: getTotalMem(){
+    return nTotalMem;
+}
 	int Node:: getFreeMem(){
 		return nFreeMem;
 	}

### 2015-05-09 src/GeoSim/Node.h

@@ -29,6 +29,8 @@ public:
 	int decreaseCPU(int cpu);
 	int increaseMem(int mem);
     int decreaseMem(int mem);
+    int getTotalCPU();
+    int getTotalMem();
 
 
 };

### 2015-05-09 src/GeoSim/common.cpp

@@ -34,6 +34,7 @@ void Barrier::Wait(int local_count)
 }
 
 void Barrier:: decrement(){
+    std::unique_lock<std::mutex> lock{_mutex};
     --_org_val;
 }
 

### 2015-05-09 src/GeoSim/main.cpp

@@ -13,10 +13,15 @@
 #include "ConfigAccessor.h"
 using namespace std;
 const string workload_chile("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
-const string workload_finland("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
-const string workload_singapore("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
-const string workload_oregon("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
-const string workload_iowa("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
+const string workload_finland("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_2.csv");
+const string workload_singapore("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_3.csv");
+const string workload_oregon("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_4.csv");
+const string workload_iowa("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_5.csv");
+const string output_chile("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/chile.trace");
+const string output_finland("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/finalnd.trace");
+const string output_singapore("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/singapore.trace");
+const string output_oregon("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/oregon.trace");
+const string output_iowa("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/iowa.trace");
 const string config_xml("config.xml");
 
 
@@ -31,11 +36,11 @@ int main()
 	
 	Barrier oBarrier(DC_COUNT);
 
-	DataCenter chile(CHILE, workload_chile,&oBarrier, "Chile", -3);
-	DataCenter finland(FINLAND, workload_finland,&oBarrier, "Finland",3);
-	DataCenter singapore(SINGAPORE, workload_singapore,&oBarrier, "Singapore", -4);
-	DataCenter oregon(OREGON, workload_oregon,&oBarrier, "Oregon", -7);
[…]

### 2015-05-09 src/GeoSim/DataCenter.cpp
Corrected Utilization prints on every update

@@ -78,11 +78,10 @@ int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
 	}
 	m_mapDCtoResource[nDCid] = mapNodes;
     string s= "Completed creating "+ name +" DataCenter at "+ getLocalTime();
-    s= "Debug utilize while creating ";
+   
     Logfile(s);
-    PrintUtilization();
-
-    Logfile(s);
+   
+   
     m_workLoad.Initialize(m_sWorkloadTrace, name, m_GMT, m_ExecutionTraces);
 
 	return SUCCESS;
@@ -107,7 +106,7 @@ void DataCenter::PrintUtilization() {
     INT64_ useMem(0);
     double pCpu;
     double pMem;
-
+    string s;
     
     for (auto node_iter = nodeMap.begin(); node_iter != nodeMap.end();
          ++node_iter) {
@@ -119,19 +118,14 @@ void DataCenter::PrintUtilization() {
         totalMem+= node_iter->second->getTotalMem();
     }
     
-    string s = name+ " Total CPU  "+ to_string(totalCpu) +"Total Memory: "+ to_string(totalMem);
-    Logfile(s);
-     s = name+ " Free CPU  "+ to_string(freeCpu) +"Free Memory: "+ to_string(freeMem);
-    Logfile(s);
-
+    
     useCpu = totalCpu - freeCpu;
     useMem = totalMem - freeMem;
-    s = name+ " Use CPU  "+ to_string(useCpu) +"Use Memory: "+ to_string(useMem);
-    Logfile(s);
+  
 
     pCpu=((double)useCpu / (double)totalCpu)*100;
     pMem=((double)useMem / (double)totalMem)*100;
-    s = name+ " CPU Utilization: "+ to_string(pCpu) +" Memory Utilization: "+ to_string(pMem);
+    s = name+ " CPU Utilization: "+ to_string(pCpu) +"% Memory Utilization: "+ to_string(pMem)+"%";
     Logfile(s);
 
     
@@ -334,9 +328,7 @@ void DataCenter::ProgressRunningJobs() {
 				m_resourceMutex.lock();
 				availResource[nodeId]->increaseCPU(Tcpu);
 				availResource[nodeId]->increaseMem(Tmem);
-                s= "Debug utilize while finished ";
-                Logfile(s);
-                PrintUtilization();
+               
                               				//removing task from node task list
[…]

### 2015-05-09 src/GeoSim/DataCenter.h

@@ -77,7 +77,6 @@ private:
 	std::list<Job*> 	GetWaitingJobs();
     string              localtime(int gmtDiff);
     string              name;
-    Log                 L;
     string              m_ExecutionTraces;
     int Logfile(string msg);
    

### 2015-05-09 src/GeoSim/Node.cpp

@@ -45,12 +45,12 @@ int Node:: getTotalMem(){
 		    return FAIL;
 		}
 	int Node::increaseMem(int mem){
-		    nFreeCPU= nFreeMem+ mem;
+		    nFreeMem= nFreeMem+ mem;
 		    return SUCCESS;
 		}
 		int Node::decreaseMem(int mem){
 			    if(nFreeMem>mem){
-			    nFreeCPU= nFreeMem- mem;
+			    nFreeMem= nFreeMem- mem;
 			    return SUCCESS;
 			    }
 			    return FAIL;

### 2015-05-09 src/GeoSim/common.cpp

@@ -38,11 +38,6 @@ void Barrier:: decrement(){
     --_org_val;
 }
 
-void Log:: print(string s){
-    m_LogMutex.lock();
-    cout<<s;
-    m_LogMutex.unlock();
-}
 
 vector<string> &Tokenize(const string &s, char delim, vector<string> &elems)
 {

### 2015-05-09 src/GeoSim/common.h

@@ -43,13 +43,6 @@ public:
 };
 
 
-class Log{
-    
-private:
-    std::mutex  m_LogMutex;
-public:
-   void print(string s);
-};
 
 
 

### 2015-05-09 src/GeoSim/DataCenter.cpp
Datacenter reads average of temp and electricity for next n hours

@@ -24,10 +24,12 @@ using namespace std;
 
 const string machineConfig("/Users/harshitdokania/Desktop/cs525/geosched/datacenters/machine_config.csv");
 
-DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string n, int gmtDiff, string tracepath) {
+DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string n, int gmtDiff, const string &tracepath, const string &tempPath, const string &elecPath) {
 	// TODO Auto-generated constructor stub
 	nDCid = id;
 	m_sWorkloadTrace = workloadPath;
+    m_TempTrace = tempPath;
+    m_ElecTrace = elecPath;
 	m_pBarrier = pBarrier;
     name = n;
     m_GMT = gmtDiff;
@@ -83,6 +85,9 @@ int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
    
    
     m_workLoad.Initialize(m_sWorkloadTrace, name, m_GMT, m_ExecutionTraces);
+    m_Temp.Initialize(m_TempTrace, name, m_GMT, m_ExecutionTraces);
+    m_Electric.Initialize(m_ElecTrace, name, m_GMT, m_ExecutionTraces);
+ 
 
 	return SUCCESS;
 }
@@ -179,6 +184,23 @@ int DataCenter:: Logfile(string msg)
     
 }
 
+
+
+double DataCenter:: TemperatureNextHours(string date, int hour){
+    
+    string hr = localhour(m_GMT);
+    int h= atoi(hr.c_str());
+    return m_Temp.TempElectricNextHours(date, hour, h);
+    
+}
+
+double DataCenter:: ElectricityNextHours(string date, int hour){
+    
+    string hr = localhour(m_GMT);
+    int h= atoi(hr.c_str());
+   return m_Electric.TempElectricNextHours(date, hour, h);
+    
+}
 
 
 
@@ -503,6 +525,7 @@ string DataCenter:: localtime(int i){
     }
     else
         Time=to_string((ptm->tm_hour+i)%24);
+     int h= atoi(Time.c_str());
     Time+=":";
     if((ptm->tm_min)<9) {
         Time+="0" ;
@@ -515,4 +538,20 @@ string DataCenter:: localtime(int i){
     Time+=to_string((ptm->tm_sec));
     return Time;
 }
[…]

### 2015-05-09 src/GeoSim/DataCenter.h

@@ -33,7 +33,7 @@ typedef std::unordered_map<int,Node*>                NodeMap;
 class Job;
 class DataCenter {
 public:
-	DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string name, int gmtDiff, string traces);
+	DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string name, int gmtDiff, const  string &traces,const string &tempPath, const string &elecPath);
 	virtual ~DataCenter();
 	int 			Initialize(DataCenterProxy * dataCenterProxies, ConfigAccessor *pAccessor);
 	void 			AddJobsToWaitingList(Job *pJob);
@@ -50,6 +50,8 @@ public:
     string          GetName();
     Barrier*        getBarrier();
     void            decrementBarrier();
+    double          TemperatureNextHours(string date, int hour);
+    double          ElectricityNextHours(string date, int hour);
 
 
 private:
@@ -63,7 +65,11 @@ private:
 	std::list<Job*>  	m_vRunningJobs;
 	std::list<Job*>  	m_vWaitingJobs;
 	GoogleTrace 		m_workLoad;
-	std::string			m_sWorkloadTrace;
+    TempElectric        m_Temp;
+    TempElectric        m_Electric;
+    std::string			m_sWorkloadTrace;
+    std::string			m_TempTrace;
+    std::string			m_ElecTrace;
 	Barrier				*m_pBarrier;
 	ConfigAccessor      *m_pAccessor;
     int                 m_GMT;
@@ -76,9 +82,11 @@ private:
 	void 				RemoveJobFromWaitingQueue(Job*);
 	std::list<Job*> 	GetWaitingJobs();
     string              localtime(int gmtDiff);
+    string              localhour(int i);
     string              name;
     string              m_ExecutionTraces;
     int Logfile(string msg);
+    
    
    
 };

### 2015-05-09 src/GeoSim/GoogleTrace.cpp

@@ -12,6 +12,7 @@
 #include <fstream>
 #include <iostream>
 #include "Job.h"
+
 using namespace std;
 
 
@@ -156,6 +157,10 @@ int GoogleTrace::Initialize(const std::string &filePath, string name, int GMT, s
 	return SUCCESS;
 }
 
+
+
+
+
 int GoogleTrace:: Logfile(string msg, string path)
 {
     
@@ -207,3 +212,115 @@ int GoogleTrace::WriteToFile(const std::string &filePath)
 	traceStream.close();
 	return SUCCESS;
 }
+traceTE::traceTE(){
+    
+}
+
+ 
+string traceTE:: parse(const std::string &sCSV){
+    vector<string> vTokens;
+    Tokenize(sCSV,',',vTokens);
+    char *pEnd;
+    if(!vTokens[YEAR].empty())
+        year= vTokens[YEAR].c_str();
+    
+    if(!vTokens[MONTH].empty())
+        month = vTokens[MONTH].c_str();
+    if(!vTokens[DAY].empty())
+        day = vTokens[DAY].c_str();
+    if(!vTokens[HOUR].empty())
+        hour = vTokens[HOUR].c_str();
+    if(!vTokens[DATA].empty())
+        TE = vTokens[DATA].c_str();
+    string date=year+month+day+hour;
+    return date;
+       
+
+    
+}
+string traceTE:: getTE(){
+    return TE;
+}
+
+int TempElectric::Initialize(const std::string &filePath, string name, int GMT, string path)
+{
+    string key, data;
+    ifstream traceStream(filePath);
+    string sLine;
+    
+    while(!traceStream.eof())
+    {
+        traceStream >> sLine;
+       // cout<<sLine;
+        
+        traceTE nDate;
+        key= nDate.parse(sLine);
+        data = nDate.getTE();
+       // cout<<"data "<<data<<endl;
+        //cout<<"key "<<key<<endl;
+        vTempElectric[key]= stoi(data);
+    }
+ 
+    
+    traceStream.close();
+    string s= name+ " Completed Temp/Electric tracing at "+ localtime(GMT);
+   Logfile(s,path);
+    return SUCCESS;
+}
+int TempElectric:: Logfile(string msg, string path)
+{
+    
+    std::fstream execTraces;
+    execTraces.open (path, std::fstream::in | std::fstream::out | std::fstream::app);
+    
+    execTraces << msg<<endl;
+    
+    execTraces.close();
+    
+    
+    return SUCCESS;
+}
[…]

### 2015-05-09 src/GeoSim/GoogleTrace.h

@@ -3,6 +3,7 @@
 
 #include "common.h"
 #include "Job.h"
+#include <unordered_map>
 #define JOB_ID  		0
 #define JOB_ARRIVAL  	1
 #define JOB_SCHEDTIME  	2
@@ -15,6 +16,12 @@
 #define JOB_SLA1	  	9
 #define JOB_SLA2	  	10
 #define JOB_SLA3	  	11
+
+#define YEAR 0
+#define MONTH 1
+#define DAY 2
+#define HOUR 3
+#define DATA 4
 
 
 struct TraceItem
@@ -48,6 +55,28 @@ struct TraceItem
 
 };
 
+
+struct traceTE{
+    string year;
+    string month;
+    string day;
+    string hour;
+    string TE;
+    traceTE();
+    string parse(const std::string &sCSV);
+    
+    string getTE();
+    
+    
+};
+
+
+
+
+
+
+
+
 class GoogleTrace
 {
 private:
@@ -61,6 +90,19 @@ public:
 	int WriteToFile(const std::string &filePath);
     int Logfile(string msg, string path);
     };
+
+class TempElectric{
+  private:
+     std::map<string,int> vTempElectric;
+    std::vector<traceTE> traces;
+  public:
+    int Initialize(const std::string &filePath, string name, int GMT, string path);
+    double TempElectricNextHours(string date, int hour, int start);
+    int  Logfile(string msg, string path);
+    
+    
+};
+
 
 
 

### 2015-05-09 src/GeoSim/Job.cpp

@@ -20,8 +20,8 @@ Job::Job(INT64_ id, INT64_ runtime, int sClass, int tasks, double cpu, double me
    nSchedClass = sClass;
    nTotalRunTime = runtime;
    nCurrRunTime = 0;
-   double myCpu = cpu/tasks;
-   double myMem = mem/tasks;
+   int myCpu = (cpu-1)/tasks +1;
+   int myMem = (mem-1)/tasks +1;
    for(int i=0; i<tasks ; i++){
 	   Task *t= new Task();
 	   t->setMem(myMem);

### 2015-05-09 src/GeoSim/common.h

@@ -11,7 +11,7 @@
 #include <string>
 #include <vector>
 #include <sstream>
-#include <cstdlib>
+
 #include <mutex>
 #include <thread>
 #include  <condition_variable>

### 2015-05-09 src/GeoSim/main.cpp

@@ -20,8 +20,20 @@ const string workload_iowa("/Users/harshitdokania/Desktop/cs525/geosched/workloa
 const string output_chile("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/chile.trace");
 const string output_finland("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/finalnd.trace");
 const string output_singapore("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/singapore.trace");
+
 const string output_oregon("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/oregon.trace");
 const string output_iowa("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/iowa.trace");
+const string temp_chile("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/chile.csv");
+const string temp_finland("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/finland.csv");
+const string temp_iowa("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/iowa.csv");
+const string temp_oregon("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/oregon.csv");
+const string temp_singapore("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/singapore.csv");
+const string elec_chile("/Users/harshitdokania/Desktop/geosched/datacenters/electricity/chile.csv");
+const string elec_finland("/Users/harshitdokania/Desktop/geosched/datacenters/electricity/finland.csv");
+const string elec_iowa("/Users/harshitdokania/Desktop/geosched/datacenters/electricity/iowa.csv");
+const string elec_oregon("/Users/harshitdokania/Desktop/geosched/datacenters/electricity/oregon.csv");
+const string elec_singapore("/Users/harshitdokania/Desktop/geosched/datacenters/electricity/singapore.csv");
+
 const string config_xml("config.xml");
 
 
@@ -36,11 +48,11 @@ int main()
 	
 	Barrier oBarrier(DC_COUNT);
 
-	DataCenter chile(CHILE, workload_chile,&oBarrier, "Chile", -3, output_chile);
[…]

### 2015-05-09 src/GeoSim/DataCenterProxy.cpp
added temp and electricity functions to proxy

@@ -40,6 +40,7 @@ string DataCenterProxy::GetName()
     
 }
 
+
 void DataCenterProxy::InformLeaving(){
     
     m_dc->decrementBarrier();
@@ -51,4 +52,12 @@ std::unordered_map<int,Node*> DataCenterProxy:: GetResourceData()
 {
 	return m_dc->GetResourceData();
 	//return std::unordered_map<int,Node*>();
+}
+
+
+double DataCenterProxy:: TemperatureNextHours(std::string date, int hour){
+    return m_dc->TemperatureNextHours(date, hour);
+}
+double DataCenterProxy::  ElectricityNextHours(std::string date, int hour){
+     return m_dc->ElectricityNextHours(date, hour);
 }

### 2015-05-09 src/GeoSim/DataCenterProxy.h

@@ -25,6 +25,8 @@ public:
 	std::unordered_map<int,Node*> GetResourceData();
     void InformLeaving();
     std::string GetName();
+    double          TemperatureNextHours(std::string date, int hour);
+    double          ElectricityNextHours(std::string date, int hour);
 private:
 	DataCenter *m_dc;
 };

### 2015-05-09 src/GeoSim/GoogleTrace.cpp
average calculation corrected to modify upto month while inreasing hours

@@ -10,6 +10,7 @@
 #include "common.h"
 #include <string>
 #include <fstream>
+#include <sstream>
 #include <iostream>
 #include "Job.h"
 
@@ -282,16 +283,22 @@ int TempElectric:: Logfile(string msg, string path)
 }
 double TempElectric:: TempElectricNextHours(string date, int hour, int start){
     std::map<string,int>:: iterator currTE;
-   // cout<<"from local  "<< date<<endl;
+   
     
     int count = 0;
     int total = 0;
     string tempDate;
     double avgTE;
     
+   
+    vector<string> vTokens;
+    Tokenize(date,',',vTokens);
+
+    
+    
     for(int i=start; (i<=23)&&(i<=start+hour);i++)
     {
-        tempDate = date + to_string(i);
+        tempDate = vTokens[0] + vTokens[1]+ vTokens[2] + to_string(i);
         currTE=vTempElectric.find(tempDate);
         if(currTE!=vTempElectric.end()){
           total+= currTE->second;
@@ -299,27 +306,49 @@ double TempElectric:: TempElectricNextHours(string date, int hour, int start){
         }
     }
     if(count!=hour){
-        char s=date[(date.size())-1];
-        int a= atoi(&s)+1;
-        std::string temp = std::to_string(a);
-        date[(date.size())-1]=temp[0];
-        int rem=hour-count;
-        for(int i=0;i<rem;i++){
+      
+        string day= vTokens[2];
+        int d= atoi(day .c_str()) +1;
+        
+        if(d<=31){
+          std::string temp = std::to_string(d);
+          date[(date.size())-1]=temp[0];
+          int rem=hour-count;
+          for(int i=0;i<rem;i++){
             
-            tempDate = date + to_string(i);
-            currTE=vTempElectric.find(tempDate);
-            if(currTE!=vTempElectric.end()){
+             tempDate = vTokens[0] + vTokens[1]+ temp + to_string(i);
+             currTE=vTempElectric.find(tempDate);
+             if(currTE!=vTempElectric.end()){
                 total+= currTE->second;
                 count++;
-            }
+             }
             
+          }
         }
-     }
-    avgTE = (double)total/(double)hour;
-    if(count==hour)
-     return avgTE;
-    else
-        return FAIL;
+        else{
+          day="1";
+          string mo = vTokens[1];
+          int  m = atoi(mo.c_str()) +1;
+          std::string temp = std::to_string(m);
+          if(m<=12){
+            int rem=hour-count;
[…]

### 2015-05-09 src/GeoSim/main.cpp

@@ -83,13 +83,13 @@ int main()
 	singapore.Initialize(dcProxy,&oAccessor);
 	oregon.Initialize(dcProxy,&oAccessor);
     iowa.Initialize(dcProxy,&oAccessor);
-   /*
-    temp and electricity
-    cout<<"Chile Temp: "<<chile.TemperatureNextHours("201313",17)<<endl;
-    cout<<"Finland Temp"<<finland.TemperatureNextHours("201313",5)<<endl;
-    cout<<"Chile.Elec:= "<<chile.ElectricityNextHours("201313",5)<<endl;
-    cout<<"Chile.Elec:= "<<chile.ElectricityNextHours("201313",5)<<endl;
-    */
+   
+   /* temp and electricity*/
+    cout<<"Chile Temp: "<<chile.TemperatureNextHours("2013,1,31",17)<<endl;
+    cout<<"Finland Temp"<<finland.TemperatureNextHours("2013,1,3",5)<<endl;
+    cout<<"Chile.Elec:= "<<chile.ElectricityNextHours("2013,1,3",5)<<endl;
+    cout<<"Chile.Elec:= "<<chile.ElectricityNextHours("2013,1,3",5)<<endl;
+    
     
     
     chile.UpdateResourceData();

### 2015-05-09 src/GeoSim/DataCenter.cpp
energ model + date time in dc

@@ -17,14 +17,23 @@
 #include "ConfigAccessor.h"
 #include <stdio.h>
 
+#define CORE_IDLE  				245
+#define CORE_PEAK  				320
+#define AVG_UTIL    			40
+#define FREE_AIR_CUT_OFF		64
+#define COP 					2 //for 15 degree celcius
+#define MONTH					1
+
 using namespace std;
 
-#define TIMEINC 				5000000   // 5 secs in us
-#define RESOURCE_SYNC_TIME		900000000 // 900 secs in us
+#define TIMEINC 				300000000   // 5 mins in us
+#define RESOURCE_SYNC_TIME		1800000000 // 1800 secs in us
+#define ENERGY_CALC_INTERVAL	900000000 // 15 mins
 
 const string machineConfig("/Users/harshitdokania/Desktop/cs525/geosched/datacenters/machine_config.csv");
 
-DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string n, int gmtDiff, const string &tracepath, const string &tempPath, const string &elecPath) {
+DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string n, int gmtDiff, const string &tracepath, \
+		const string &tempPath, const string &elecPath, bool isAirEco) {
 	// TODO Auto-generated constructor stub
 	nDCid = id;
 	m_sWorkloadTrace = workloadPath;
@@ -34,8 +43,10 @@ DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrie
     name = n;
     m_GMT = gmtDiff;
     m_ExecutionTraces = tracepath;
-
-    
+    bAirEco = isAirEco;
+    m_TotalCost = 0;
+    m_TotalEnergy = 0;
+    InitStartPoint();
     string s= "Started creating " + name + " DataCenter at " +getLocalTime();
    
     Logfile(s);
@@ -43,6 +54,11 @@ DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrie
 
 DataCenter::~DataCenter() {
 	// TODO Auto-generated destructor stub
+}
+
+double 	DataCenter::isAirEco()
+{
+	return bAirEco;
 }
 
 int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
[…]

### 2015-05-09 src/GeoSim/DataCenter.h

@@ -17,6 +17,9 @@
 #include <mutex>
 #include "GoogleTrace.h"
 #include "common.h"
+#include <chrono>
+#include <ctime>
+
 class DataCenterProxy; //forward declare
 class Node;
 class ConfigAccessor;
@@ -33,12 +36,13 @@ typedef std::unordered_map<int,Node*>                NodeMap;
 class Job;
 class DataCenter {
 public:
-	DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string name, int gmtDiff, const  string &traces,const string &tempPath, const string &elecPath);
+	DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string name, int gmtDiff, const  string &traces\
+			,const string &tempPath, const string &elecPath, bool);
 	virtual ~DataCenter();
 	int 			Initialize(DataCenterProxy * dataCenterProxies, ConfigAccessor *pAccessor);
 	void 			AddJobsToWaitingList(Job *pJob);
 	void 			ScheduleJobsFromWaitingList();
-	void 			PrintUtilization();
+	void 			GetUtilization(double &util, int &free, int &total_cpu);
 	NodeMap 		GetResourceData();
 	void 			UpdateResourceData();
 	void 			set_dataCenterProxies(DataCenterProxy *proxy);
@@ -50,9 +54,9 @@ public:
     string          GetName();
     Barrier*        getBarrier();
     void            decrementBarrier();
-    double          TemperatureNextHours(string date, int hour);
-    double          ElectricityNextHours(string date, int hour);
-
+    double          TemperatureNextHours(int hour);
+    double          ElectricityNextHours(int hour);
+    double 			isAirEco();
 
 private:
 	int 				nDCid;
@@ -73,7 +77,13 @@ private:
 	Barrier				*m_pBarrier;
 	ConfigAccessor      *m_pAccessor;
     int                 m_GMT;
-	bool				CheckDCFit(Job *,NodeMap &);
+    bool				bAirEco;
+    double 				m_TotalCost;
+    double				m_TotalEnergy;
+    std::chrono::system_clock::time_point startPoint;
+    INT64_ 				arrivalTime;
+private:
+    bool				CheckDCFit(Job *,NodeMap &);
 	void 				StartSimulation(); //main scheduling logic goes in here
 	std::vector<int>    GetDCSchedulable(Job *);
 	void				MetaSchedJobToDC(std::vector<int> vecDCs,Job*);
@@ -85,7 +95,14 @@ private:
[…]

### 2015-05-09 src/GeoSim/DataCenterProxy.cpp

@@ -46,7 +46,10 @@ void DataCenterProxy::InformLeaving(){
     m_dc->decrementBarrier();
 }
 
-
+bool			DataCenterProxy::isAirEco()
+{
+	return m_dc->isAirEco();
+}
 
 std::unordered_map<int,Node*> DataCenterProxy:: GetResourceData()
 {
@@ -55,9 +58,9 @@ std::unordered_map<int,Node*> DataCenterProxy:: GetResourceData()
 }
 
 
-double DataCenterProxy:: TemperatureNextHours(std::string date, int hour){
-    return m_dc->TemperatureNextHours(date, hour);
+double DataCenterProxy:: TemperatureNextHours(int hour){
+    return m_dc->TemperatureNextHours(hour);
 }
-double DataCenterProxy::  ElectricityNextHours(std::string date, int hour){
-     return m_dc->ElectricityNextHours(date, hour);
+double DataCenterProxy::  ElectricityNextHours(int hour){
+     return m_dc->ElectricityNextHours( hour);
 }

### 2015-05-09 src/GeoSim/DataCenterProxy.h

@@ -25,8 +25,9 @@ public:
 	std::unordered_map<int,Node*> GetResourceData();
     void InformLeaving();
     std::string GetName();
-    double          TemperatureNextHours(std::string date, int hour);
-    double          ElectricityNextHours(std::string date, int hour);
+    double          TemperatureNextHours(int hour);
+    double          ElectricityNextHours(int hour);
+    bool			isAirEco();
 private:
 	DataCenter *m_dc;
 };

### 2015-05-09 src/GeoSim/Job.cpp

@@ -20,6 +20,7 @@ Job::Job(INT64_ id, INT64_ runtime, int sClass, int tasks, double cpu, double me
    nSchedClass = sClass;
    nTotalRunTime = runtime;
    nCurrRunTime = 0;
+   nCores = cpu;
    int myCpu = (cpu-1)/tasks +1;
    int myMem = (mem-1)/tasks +1;
    for(int i=0; i<tasks ; i++){

### 2015-05-09 src/GeoSim/Job.h

@@ -53,6 +53,7 @@ public:
 	std::vector<Task*> getTasks();
 	INT64_ GetTotalRunTime();
     // add required method
+	double   nCores;
 private:
 	std::vector<Task*> vTasks;
 	INT64_ nTotalRunTime;
@@ -60,6 +61,7 @@ private:
 	INT64_ nJobID;
     int nodeID;
 	int    nSchedClass;
+	 //neeed for energy calc
 };
 
 #endif /* JOB_H_ */

### 2015-05-09 src/GeoSim/main.cpp

@@ -12,27 +12,31 @@
 #include "DataCenterProxy.h"
 #include "ConfigAccessor.h"
 using namespace std;
-const string workload_chile("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_1.csv");
-const string workload_finland("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_2.csv");
-const string workload_singapore("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_3.csv");
-const string workload_oregon("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_4.csv");
-const string workload_iowa("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/5_5.csv");
-const string output_chile("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/chile.trace");
-const string output_finland("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/finalnd.trace");
-const string output_singapore("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/singapore.trace");
 
-const string output_oregon("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/oregon.trace");
-const string output_iowa("/Users/harshitdokania/Desktop/cs525/geosched/workloads/google/iowa.trace");
-const string temp_chile("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/chile.csv");
-const string temp_finland("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/finland.csv");
-const string temp_iowa("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/iowa.csv");
-const string temp_oregon("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/oregon.csv");
-const string temp_singapore("/Users/harshitdokania/Desktop/geosched/datacenters/temperature/singapore.csv");
-const string elec_chile("/Users/harshitdokania/Desktop/geosched/datacenters/electricity/chile.csv");
-const string elec_finland("/Users/harshitdokania/Desktop/geosched/datacenters/electricity/finland.csv");
-const string elec_iowa("/Users/harshitdokania/Desktop/geosched/datacenters/electricity/iowa.csv");
[…]

### 2015-05-09 src/GeoSim/DataCenter.cpp
added extra logs

@@ -54,6 +54,7 @@ DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrie
 
 DataCenter::~DataCenter() {
 	// TODO Auto-generated destructor stub
+	 execTraces.close();
 }
 
 double 	DataCenter::isAirEco()
@@ -103,7 +104,7 @@ int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
     m_workLoad.Initialize(m_sWorkloadTrace, name, m_GMT, m_ExecutionTraces);
     m_Temp.Initialize(m_TempTrace, name, m_GMT, m_ExecutionTraces);
     m_Electric.Initialize(m_ElecTrace, name, m_GMT, m_ExecutionTraces);
- 
+    execTraces.open (m_ExecutionTraces, std::fstream::in | std::fstream::out | std::fstream::app);
 
 	return SUCCESS;
 }
@@ -129,6 +130,7 @@ void DataCenter::GetUtilization(double &cpu_util,int &free_cpu, int &total_cpu) 
     double pMem;
     string s;
     
+    m_resourceMutex.lock();
     for (auto node_iter = nodeMap.begin(); node_iter != nodeMap.end();
          ++node_iter) {
         
@@ -138,7 +140,7 @@ void DataCenter::GetUtilization(double &cpu_util,int &free_cpu, int &total_cpu) 
         totalCpu+= node_iter->second->getTotalCPU();
         totalMem+= node_iter->second->getTotalMem();
     }
-    
+    m_resourceMutex.unlock();
     
     useCpu = totalCpu - freeCpu;
 
@@ -150,7 +152,7 @@ void DataCenter::GetUtilization(double &cpu_util,int &free_cpu, int &total_cpu) 
     cpu_util = pCpu;
     free_cpu = freeCpu;
     total_cpu = totalCpu;
-    s = name+ " CPU Utilization: "+ to_string(pCpu) +"% Memory Utilization: "+ to_string(pMem)+"%";
+    s = name + "Time: " + to_string(arrivalTime) + " CPU Utilization: "+ to_string(pCpu) +"% Memory Utilization: "+ to_string(pMem)+"%";
     Logfile(s);
 
     
@@ -191,15 +193,9 @@ GoogleTrace* DataCenter::getWorkLoad() {
 
 int DataCenter:: Logfile(string msg)
 {
-  
-    std::fstream execTraces;
[…]

### 2015-05-09 src/GeoSim/DataCenter.h

@@ -19,6 +19,7 @@
 #include "common.h"
 #include <chrono>
 #include <ctime>
+#include <fstream>
 
 class DataCenterProxy; //forward declare
 class Node;
@@ -82,6 +83,7 @@ private:
     double				m_TotalEnergy;
     std::chrono::system_clock::time_point startPoint;
     INT64_ 				arrivalTime;
+    std::fstream 		execTraces;
 private:
     bool				CheckDCFit(Job *,NodeMap &);
 	void 				StartSimulation(); //main scheduling logic goes in here

### 2015-05-09 src/GeoSim/GoogleTrace.cpp

@@ -346,7 +346,6 @@ double TempElectric:: TempElectricNextHours(string date, int hour, int start){
     }
     
     avgTE = (double)total/(double)count;
-    
     return avgTE;
    
     

### 2015-05-09 src/GeoSim/DataCenter.cpp
removing jobs after max time

@@ -30,6 +30,9 @@ using namespace std;
 #define RESOURCE_SYNC_TIME		1800000000 // 1800 secs in us
 #define ENERGY_CALC_INTERVAL	900000000 // 15 mins
 
+#define MAX_TIME				14400000000 //4 hours
+
+
 const string machineConfig("/Users/harshitdokania/Desktop/cs525/geosched/datacenters/machine_config.csv");
 
 DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrier, string n, int gmtDiff, const string &tracepath, \
@@ -319,7 +322,7 @@ void DataCenter::StartSimulation() {
         cycle_count++;
 	}
     
-    DecreaseBarrier();
+	decrementBarrier();
     s =name + " Leaving simulation at"+ getLocalTime() + " TotalCost:" + to_string(m_TotalCost) + \
     		" Total Energy:" + to_string(m_TotalEnergy);
     Logfile(s);
@@ -365,16 +368,6 @@ void DataCenter::MetaSchedJobToDC(std::vector<int> vecDCs, Job* pJob)
 	}
 }
 
-void DataCenter:: DecreaseBarrier(){
-    for (int i = 0; i < DC_COUNT; i++) {
-        if (i != nDCid){
-            m_dataCenterProxies[i].InformLeaving();
-            string s= name+ " Decreasing Barrier on "+ m_dataCenterProxies[i].GetName()+ " time: " +getLocalTime();
-            Logfile(s);
-        }
-    }
-}
-
 
 vector<int> DataCenter::GetDCSchedulable(Job *pJob)
 {
@@ -386,13 +379,28 @@ vector<int> DataCenter::GetDCSchedulable(Job *pJob)
 	}
 	return ret;
 }
-string DataCenter:: GetName(){
+
+
+string DataCenter:: GetName()
+{
     return name;
 }
 
 void DataCenter::ProgressRunningJobs() {
 	NodeMap &availResource = m_mapDCtoResource[nDCid];
     string s;
+    if(arrivalTime > MAX_TIME)
+	{
+    	string log = "time up. remvoing jobs from running list.";
+    	Logfile(log);
+		// delete all running jobs
+    	for(auto iter = m_vRunningJobs.begin(); iter != m_vRunningJobs.end(); ++iter)
+    	{
+    		Job *pJob = *iter;
+    		delete pJob;
[…]

### 2015-05-09 src/GeoSim/DataCenter.h

@@ -47,7 +47,6 @@ public:
 	NodeMap 		GetResourceData();
 	void 			UpdateResourceData();
 	void 			set_dataCenterProxies(DataCenterProxy *proxy);
-    void            DecreaseBarrier();
 	GoogleTrace* 	getWorkLoad();
 	void 			Join();
 	void 			Simulation();
@@ -94,7 +93,6 @@ private:
 	void 				RemoveJobFromWaitingQueue(Job*);
 	std::list<Job*> 	GetWaitingJobs();
     string              localtime(int gmtDiff);
-    string              localhour(int i);
     string              name;
     string              m_ExecutionTraces;
     int 				Logfile(string msg);

### 2015-05-09 src/GeoSim/DataCenterProxy.cpp

@@ -41,10 +41,6 @@ string DataCenterProxy::GetName()
 }
 
 
-void DataCenterProxy::InformLeaving(){
-    
-    m_dc->decrementBarrier();
-}
 
 bool			DataCenterProxy::isAirEco()
 {

### 2015-05-10 src/GeoSim/DataCenter.cpp
Corrected utilization

@@ -19,7 +19,7 @@
 
 #define CORE_IDLE  				245
 #define CORE_PEAK  				320
-#define AVG_UTIL    			40
+#define AVG_UTIL    			0.4
 #define FREE_AIR_CUT_OFF		64
 #define COP 					2 //for 15 degree celcius
 #define MONTH					1
@@ -139,16 +139,35 @@ void DataCenter::GetUtilization(double &cpu_util,int &free_cpu, int &total_cpu) 
         
 
         freeCpu+= node_iter->second->getFreeCPU();
+   #ifdef DEBUG
+        s="free CPU "+ to_string(freeCpu)+ " Node id "+ to_string(node_iter->second->getNodeID());
+        Logfile(s);
+#endif
         freeMem+= node_iter->second->getFreeMem();
+        #ifdef DEBUG
+        s="free Mem "+ to_string(freeMem)+ " Node id "+ to_string(node_iter->second->getNodeID());
+        Logfile(s);
+#endif
         totalCpu+= node_iter->second->getTotalCPU();
+        #ifdef DEBUG
+        s="free Total CPU "+ to_string(totalCpu)+ " Node id "+ to_string(node_iter->second->getNodeID());
+        Logfile(s);
+#endif
         totalMem+= node_iter->second->getTotalMem();
+        #ifdef DEBUG
+        s="free Total Mem "+ to_string(totalMem)+ " Node id "+ to_string(node_iter->second->getNodeID());
+        Logfile(s);
+        #endif
     }
     m_resourceMutex.unlock();
     
     useCpu = totalCpu - freeCpu;
 
     useMem = totalMem - freeMem;
-  
+#ifdef DEBUG
+    s=" while utilization used " + to_string(useCpu)+ "  free CPU "+ to_string(freeCpu)  + " total cpu "+to_string(totalCpu) + " at  time: " +getLocalTime();
+    Logfile(s);
+#endif
 
     pCpu=((double)useCpu / (double)totalCpu)*100;
     pMem=((double)useMem / (double)totalMem)*100;
@@ -196,14 +215,26 @@ GoogleTrace* DataCenter::getWorkLoad() {
 
 int DataCenter:: Logfile(string msg)
 {
-
+    m_printMutex.lock();
     execTraces << msg<<endl;
-
+m_printMutex.unlock();
     return SUCCESS;
     
 }
 
-
+int DataCenter:: Logfile(double value,string n)
+{
[…]

### 2015-05-10 src/GeoSim/DataCenter.h

@@ -66,6 +66,7 @@ private:
 	NodeMap				m_mapNodes; // nodeid-> node map
 	std::mutex          m_waitMutex;
 	std::mutex          m_resourceMutex;
+    std::mutex          m_printMutex;
 	std::list<Job*>  	m_vRunningJobs;
 	std::list<Job*>  	m_vWaitingJobs;
 	GoogleTrace 		m_workLoad;
@@ -96,9 +97,10 @@ private:
     string              name;
     string              m_ExecutionTraces;
     int 				Logfile(string msg);
+    int 				Logfile(double value, string n);
     // speculated cost calculation
-    double 				CalculateDynamicJobCost(DataCenterProxy* proxy,Job *pJob);
-    double 				CalculateCoolingCost(DataCenterProxy* proxy,Job *pJob);
+    double 				CalculateDynamicJobCost(DataCenterProxy* proxy,Job *pJob, string name);
+    double 				CalculateCoolingCost(DataCenterProxy* proxy,Job *pJob, string name);
     //
     void 				UpdateEnergyCost();
     void 				InitStartPoint();

### 2015-05-10 src/GeoSim/GoogleTrace.h

@@ -34,8 +34,8 @@ struct TraceItem
 	INT64_ 			nFinishTime;   // job event
 	int			    nSchedClass; // job event, may not need it once SLA is decided
 	int 			nTasks;
-	double			fTotalCPU;
-	double 			fTotalMem;
+    int 			fTotalCPU;
+	int 			fTotalMem;
 	std::string 	sUser;      // job event
 	double 			fSLAParam1; // synthetic
 	double 			fSLAParam2; // synthetic

### 2015-05-10 src/GeoSim/Job.cpp

@@ -14,15 +14,15 @@ Job::Job()
 {
 }
 
-Job::Job(INT64_ id, INT64_ runtime, int sClass, int tasks, double cpu, double mem) {
+Job::Job(INT64_ id, INT64_ runtime, int sClass, int tasks, int cpu, int mem) {
 	// TODO Auto-generated constructor stub
    nJobID = id;
    nSchedClass = sClass;
    nTotalRunTime = runtime;
    nCurrRunTime = 0;
    nCores = cpu;
-   int myCpu = (cpu-1)/tasks +1;
-   int myMem = (mem-1)/tasks +1;
+   int myCpu = (cpu)/tasks;
+   int myMem = (mem)/tasks;
    for(int i=0; i<tasks ; i++){
 	   Task *t= new Task();
 	   t->setMem(myMem);

### 2015-05-10 src/GeoSim/Job.h

@@ -42,7 +42,7 @@ private:
 class Job {
 public:
     Job();
-	Job(INT64_, INT64_, int,  int , double , double);
+	Job(INT64_, INT64_, int,  int , int, int);
 	virtual ~Job();
 	INT64_ getJobID();
     int setNodeID(int n);
@@ -53,7 +53,7 @@ public:
 	std::vector<Task*> getTasks();
 	INT64_ GetTotalRunTime();
     // add required method
-	double   nCores;
+	int   nCores;
 private:
 	std::vector<Task*> vTasks;
 	INT64_ nTotalRunTime;

### 2015-05-10 src/GeoSim/Node.cpp

@@ -34,11 +34,12 @@ int Node:: getTotalMem(){
 		return nNodeID;
 	}
 	int Node::increaseCPU(int cpu){
+       
 	    nFreeCPU= nFreeCPU+ cpu;
 	    return SUCCESS;
 	}
 	int Node::decreaseCPU(int cpu){
-		    if(nFreeCPU>cpu){
+		    if(nFreeCPU>=cpu){
 		    nFreeCPU= nFreeCPU- cpu;
 		    return SUCCESS;
 		    }
@@ -49,7 +50,7 @@ int Node:: getTotalMem(){
 		    return SUCCESS;
 		}
 		int Node::decreaseMem(int mem){
-			    if(nFreeMem>mem){
+			    if(nFreeMem>=mem){
 			    nFreeMem= nFreeMem- mem;
 			    return SUCCESS;
 			    }

### 2015-05-10 src/GeoSim/DataCenter.cpp
minor decalaration bug

@@ -370,6 +370,7 @@ void DataCenter::  decrementBarrier(){
 
 void DataCenter::MetaSchedJobToDC(std::vector<int> vecDCs, Job* pJob)
 {
+        string s;
 	if(vecDCs.size() == 0)
 	{
 		//skip the job
@@ -389,7 +390,7 @@ void DataCenter::MetaSchedJobToDC(std::vector<int> vecDCs, Job* pJob)
 		double total_cost = CalculateDynamicJobCost(&m_dataCenterProxies[vecDCs[0]],pJob, m_dataCenterProxies[0].GetName()) + \
 				CalculateCoolingCost(&m_dataCenterProxies[vecDCs[0]],pJob,m_dataCenterProxies[0].GetName());
         #ifdef DEBUG
-        string s = name + " recieved total cost from " + m_dataCenterProxies[0].GetName();
+         s = name + " recieved total cost from " + m_dataCenterProxies[0].GetName();
         Logfile(total_cost, s );
 #endif
 		for (int index = 1; index < vecDCs.size(); ++index)

### 2015-05-10 src/GeoSim/DataCenter.cpp
Cool load balancer and energy calculation fixed

@@ -40,24 +40,24 @@ DataCenter::DataCenter(int id, const std::string &workloadPath, Barrier *pBarrie
 	// TODO Auto-generated constructor stub
 	nDCid = id;
 	m_sWorkloadTrace = workloadPath;
-    m_TempTrace = tempPath;
-    m_ElecTrace = elecPath;
+	m_TempTrace = tempPath;
+	m_ElecTrace = elecPath;
 	m_pBarrier = pBarrier;
-    name = n;
-    m_GMT = gmtDiff;
-    m_ExecutionTraces = tracepath;
-    bAirEco = isAirEco;
-    m_TotalCost = 0;
-    m_TotalEnergy = 0;
-    InitStartPoint();
-    string s= "Started creating " + name + " DataCenter at " +getLocalTime();
-   
-    Logfile(s);
-   }
+	name = n;
+	m_GMT = gmtDiff;
+	m_ExecutionTraces = tracepath;
+	bAirEco = isAirEco;
+	m_TotalCost = 0;
+	m_TotalEnergy = 0;
+	InitStartPoint();
+	string s= "Started creating " + name + " DataCenter at " +getLocalTime();
+
+	Logfile(s);
+}
 
 DataCenter::~DataCenter() {
 	// TODO Auto-generated destructor stub
-	 execTraces.close();
+	execTraces.close();
 }
 
 double 	DataCenter::isAirEco()
@@ -76,7 +76,7 @@ int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
 	int memAvail;
 	ifstream fileStream(machineConfig);
 	int nodeID = 0;
-   
+
 
 	NodeMap mapNodes;
 	while (!fileStream.eof()) {
@@ -99,85 +99,159 @@ int DataCenter::Initialize(DataCenterProxy * dataCenterProxies,
 		}
 	}
 	m_mapDCtoResource[nDCid] = mapNodes;
-    string s= "Completed creating "+ name +" DataCenter at "+ getLocalTime();
-   
-    Logfile(s);
-   
-   
-    m_workLoad.Initialize(m_sWorkloadTrace, name, m_GMT, m_ExecutionTraces);
-    m_Temp.Initialize(m_TempTrace, name, m_GMT, m_ExecutionTraces);
-    m_Electric.Initialize(m_ElecTrace, name, m_GMT, m_ExecutionTraces);
-    execTraces.open (m_ExecutionTraces, std::fstream::in | std::fstream::out | std::fstream::app);
[…]

### 2015-05-10 src/GeoSim/DataCenter.h

@@ -57,6 +57,7 @@ public:
     double          TemperatureNextHours(int hour);
     double          ElectricityNextHours(int hour);
     double 			isAirEco();
+    double          GetUtilization();
 
 private:
 	int 				nDCid;
@@ -101,6 +102,7 @@ private:
     // speculated cost calculation
     double 				CalculateDynamicJobCost(DataCenterProxy* proxy,Job *pJob, string name);
     double 				CalculateCoolingCost(DataCenterProxy* proxy,Job *pJob, string name);
+    double              CoolLoadBalancer(DataCenterProxy* proxy,Job *pJob, string n);
     //
     void 				UpdateEnergyCost();
     void 				InitStartPoint();

### 2015-05-10 src/GeoSim/DataCenterProxy.cpp

@@ -60,3 +60,11 @@ double DataCenterProxy:: TemperatureNextHours(int hour){
 double DataCenterProxy::  ElectricityNextHours(int hour){
      return m_dc->ElectricityNextHours( hour);
 }
+double   DataCenterProxy::       GetUtilization(){
+    return m_dc->GetUtilization();
+}
+
+
+
+
+

### 2015-05-10 src/GeoSim/DataCenterProxy.h

@@ -27,8 +27,10 @@ public:
     double          TemperatureNextHours(int hour);
     double          ElectricityNextHours(int hour);
     bool			isAirEco();
+    double          GetUtilization();
 private:
 	DataCenter *m_dc;
+    
 };
 
 #endif /* DATACENTERPROXY_H_ */

### 2015-05-10 datacenters/electricity.py
1 day trace

@@ -10,7 +10,7 @@ import calendar
 
 cwd=os.getcwd()
 
-address="/Users/harshitdokania/Desktop/cs525/geosched/datacenters/electricity"
+address="/Users/anirudhnair/code/GeoSched/datacenters/electricity"
 
 
 

### 2015-05-10 src/GeoSim/GoogleTrace.cpp
resolved macro conflicts

@@ -225,8 +225,8 @@ string traceTE:: parse(const std::string &sCSV){
     if(!vTokens[YEAR].empty())
         year= vTokens[YEAR].c_str();
     
-    if(!vTokens[MONTH].empty())
-        month = vTokens[MONTH].c_str();
+    if(!vTokens[MONT].empty())
+        month = vTokens[MONT].c_str();
     if(!vTokens[DAY].empty())
         day = vTokens[DAY].c_str();
     if(!vTokens[HOUR].empty())

### 2015-05-10 src/GeoSim/GoogleTrace.h

@@ -18,7 +18,7 @@
 #define JOB_SLA3	  	11
 
 #define YEAR 0
-#define MONTH 1
+#define MONT 1
 #define DAY 2
 #define HOUR 3
 #define DATA 4

### 2015-05-10 src/GeoSim/DataCenter.cpp
traces for 1024 nodes, each with 800 cores

@@ -30,7 +30,7 @@ using namespace std;
 #define RESOURCE_SYNC_TIME		1800000000 // 1800 secs in us
 #define ENERGY_CALC_INTERVAL	900000000 // 15 mins
 
-#define MAX_TIME				14400000000 //4 hours
+#define MAX_TIME				86400000000 //24 hours
 
 
 const string machineConfig("../../datacenters/machine_config.csv");

### 2015-05-10 src/GeoSim/main.cpp

@@ -14,11 +14,11 @@
 using namespace std;
 
 //all these paths are relative
-const string workload_chile("../../workloads/google/5_1_.csv");
-const string workload_finland("../../workloads/google/5_2_.csv");
-const string workload_singapore("../../workloads/google/5_3_.csv");
-const string workload_oregon("../../workloads/google/5_4_.csv");
-const string workload_iowa("../../workloads/google/5_5_.csv");
+const string workload_chile("../../workloads/google/5_1_1day.csv");
+const string workload_finland("../../workloads/google/5_2_1day.csv");
+const string workload_singapore("../../workloads/google/5_3_1day.csv");
+const string workload_oregon("../../workloads/google/5_4_1day.csv");
+const string workload_iowa("../../workloads/google/5_5_1day.csv");
 
 const string output_chile("chile.trace");
 const string output_finland("finalnd.trace");

### 2015-05-13 scheduler/CloudSimExample9.java
cleaning for public release

@@ -1,315 +1,0 @@
-
-package org.cloudbus.cloudsim.examples;
-
-/* @author harshitdokania */
-
-import java.text.DecimalFormat;
-import java.util.ArrayList;
-import java.util.Calendar;
-import java.util.LinkedList;
-import java.util.List;
-
-import org.cloudbus.cloudsim.Cloudlet;
-import org.cloudbus.cloudsim.CloudletSchedulerTimeShared;
-import org.cloudbus.cloudsim.Datacenter;
-import org.cloudbus.cloudsim.DatacenterBroker;
-import org.cloudbus.cloudsim.DatacenterCharacteristics;
-import org.cloudbus.cloudsim.Host;
-import org.cloudbus.cloudsim.Log;
-import org.cloudbus.cloudsim.Pe;
-import org.cloudbus.cloudsim.Storage;
-import org.cloudbus.cloudsim.Vm;
-import org.cloudbus.cloudsim.VmAllocationPolicySimple;
-import org.cloudbus.cloudsim.VmSchedulerTimeShared;
-import org.cloudbus.cloudsim.core.CloudSim;
-import org.cloudbus.cloudsim.provisioners.BwProvisionerSimple;
-import org.cloudbus.cloudsim.provisioners.PeProvisionerSimple;
-import org.cloudbus.cloudsim.provisioners.RamProvisionerSimple;
-import org.cloudbus.cloudsim.util.WorkloadFileReader;
-
-/**
- * A simple example showing how to create a datacenter with one host and run one
- * cloudlet on it.
- */
-public class CloudSimeExample9 {
-
-    /** The cloudlet list. */
-    private static List<Cloudlet> cloudletList;
-
-    /** The vmlist. */
-    private static List<Vm> vmlist;
-
-    /**
-     * Creates main() to run this example.
-     * 
-     * @param args
-     *        the args
-     *        
-     */
-    
-    private static List<Vm> createVM(int userId, int vms) {
-		//Creates a container to store VMs. This list is passed to the broker later
-		LinkedList<Vm> list = new LinkedList<Vm>();
-
-		//VM Parameters
-		long size = 10000; //image size (MB)
-		int ram = 512; //vm memory (MB)
-		int mips = 1000;
-		long bw = 1000;
-		int pesNumber = 1; //number of cpus
-		String vmm = "Xen"; //VMM name
-
-		//create VMs
-		Vm[] vm = new Vm[vms];
-
[…]

### 2015-05-13 scheduler/WorkloadFileReader.java

@@ -1,568 +1,0 @@
-/*
- * Title:        CloudSim Toolkit
- * Description:  CloudSim (Cloud Simulation) Toolkit for Modeling and Simulation of Clouds
- * Licence:      GPL - http://www.gnu.org/copyleft/gpl.html
- *
- * Copyright (c) 2009-2012, The University of Melbourne, Australia
- */
-
-package org.cloudbus.cloudsim.util;
-
-import java.io.BufferedReader;
-import java.io.File;
-import java.io.FileInputStream;
-import java.io.FileNotFoundException;
-import java.io.IOException;
-import java.io.InputStreamReader;
-import java.util.ArrayList;
-import java.util.Enumeration;
-import java.util.zip.GZIPInputStream;
-import java.util.zip.ZipEntry;
-import java.util.zip.ZipFile;
-
-import org.cloudbus.cloudsim.Cloudlet;
-import org.cloudbus.cloudsim.UtilizationModel;
-import org.cloudbus.cloudsim.UtilizationModelFull;
-
-/**
- * This class is responsible for reading resource traces from a file and creating a list of jobs.
- * <p>
- * <b>NOTE:</b>
- * <ul>
- * <li>This class can only take <tt>one</tt> trace file of the following format: <i>ASCII text, zip,
- * gz.</i>
- * <li>If you need to load multiple trace files, then you need to create multiple instances of this
- * class <tt>each with a unique 
- *      	 entity name</tt>.
- * <li>If size of the trace file is huge or contains lots of traces please increase the JVM heap
- * size accordingly by using <tt>java -Xmx</tt> option when running the simulation.
- * <li>The default job file size for sending to and receiving from a resource is
- * {@link gridsim.net.Link#DEFAULT_MTU}. However, you can specify the file size by using
- * {@link #setGridletFileSize(int)}.
- * <li>A job run time is only for 1 PE <tt>not</tt> the total number of allocated PEs. Therefore, a
- * Gridlet length is also calculated for 1 PE.<br>
- * For example, job #1 in the trace has a run time of 100 seconds for 2 processors. This means each
- * processor runs job #1 for 100 seconds, if the processors have the same specification.
- * </ul>
- * <p>
- * By default, this class follows the standard workload format as specified in <a
- * href="http://www.cs.huji.ac.il/labs/parallel/workload/">
[…]

### 2015-05-13 doc/report/final/main.tex
final report

\documentclass[conference,12pt]{IEEEtran}
\usepackage{blindtext, graphicx}
\usepackage[]{graphicx}
\usepackage[sort, numbers]{natbib}
\usepackage{hyperref}
\usepackage{adjustbox,lipsum}
\usepackage{subfig}
\usepackage{float}
\usepackage{amssymb,amsmath}
\usepackage{textcomp}

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Anirudh Jayakumar}
\IEEEauthorblockA{Department of Computer Science\\
University of Illinois at Urbana-Champaign\\
Email: ajayaku2@illinois.edu}
\and
\IEEEauthorblockN{Harshit Dokania}
\IEEEauthorblockA{Department of Computer Science\\
University of Illinois at Urbana-Champaign\\
Email: hdokani2@illinois.edu}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

\newenvironment{list1}{
  \begin{list}{\ding{113}}{%
      \setlength{\itemsep}{0in}
      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
      \setlength{\topsep}{0in} \setlength{\partopsep}{0in}
      \setlength{\leftmargin}{0.17in}}}{\end{list}}
\newenvironment{list2}{
  \begin{list}{$\bullet$}{%
      \setlength{\itemsep}{0in}
      \setlength{\parsep}{0in} \setlength{\parskip}{0in}
      \setlength{\topsep}{0in} \setlength{\partopsep}{0in}
      \setlength{\leftmargin}{0.2in}}}{\end{list}}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
Today, many of the leading cloud service providers have geographically distributed data centers to serve millions of users around the world. The need for a geo-distributed data center arises from the need to provide low-latency and  high-availability of the different services. The workloads that run on these data centers are also highly diverse. A typical data centers run jobs of different kinds including business-critical workloads---web servers, mail servers, instant messaging services etc, big-data processing, real-time data analytics and HPC jobs. As cloud services grow to serve more customers, there is an increasing need for a workload provisioning mechanism which can minimize the cost of operation of the data center while minimizing SLA violations and also minimizing the user-perceived latency.

We present GeoSched, a job provisioning framework that is workload aware, energy-aware and cooling-aware. GeoSched considers the latency sensitivity of the incoming jobs, electricity pricing of the region and cooling techniques used in the data center to schedule jobs on different data centers. To evaluate our scheme, we perform extensive simulations on real-world workload traces.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Cloud Computing, Energy consumption, Resource management, Data center.
\end{IEEEkeywords}

\section{INTRODUCTION}
\label{sec:intro}
The emergence of cloud computing has resulted in an array of services being provided over the Internet. These services include office applications, raw compute power, storage and a variety of Internet services including social media, multimedia streaming among many others. Also due to the explosion of digital content, big data, e-commerce, and Internet traffic, there is a continuous effort to expand the capacity of data centers. This demand has increased the need for efficient data centers to handle the large amount computing resource requests especially among Internet companies like Facebook, Google, Microsoft and Amazon. The size of these data centers can be more than 100,000 servers \cite{greenberg2008cost} resulting in a huge costs, both to procure the IT infrastructure and to keep the systems running.

A typical data center consists of three major subsystems: servers and network cables form the IT equipments, power subsystem that supports the data center and the cooling equipments that help to maintain the temperature of the data center for the IT equipments to work effectively. The amount of energy required to smoothly run a data center is extremely high. In 2013 alone, U.S data centers have consumed 91 billion kilowatt-hours of electricity \cite{nrdcreport}, which is the amount of electricity needed to power New York City twice over. This energy consumption is expected to reach 140 billion kilowatt-hours by 2020. A major part of the energy consumption is due to cooling energy, the energy required to maintain a specific temperature in the data center. According to some reports the cooling energy can be as high as 50\% \cite{sullivan2002alternating} \cite{patel2003smart} \cite{sawyer2004calculating} of the total energy budget.

In this paper we propose a meta-scheduler which schedules jobs to those data center which will cost the least to run an incoming job. This will in-turn reduce the total cost of operations and can benefit enterprises that operate geo-distributed data centers. Our scheduler take into account the server running costs as well as the cooling costs. The cooling costs depends on the kind of equipment used. In a traditional data center, chillers are used to cool down the hot air from CRAC units using mechanical refrigeration. This mechanism incurs large amount of power \cite{zhou2012optimization}. Recently, most data centers have adopted very cost effective mechanism called air-side economizer to cool the servers. An air-side economizer pulls the outside air into the data center and the hot air is directed outside, thus cooling the IT equipments in the data center. The use of such economizer is restricted to cooler regions. We exploit the fact that geo-distributed data centers will have different temperatures and hence the cooling efficiency of data centers will differ. 
%Last line

The rest of the paper is organized as follows. Section \ref{sec:related} presents the related work on cost minimization of data centers and other energy-efficiency research in the cloud computing. Section \ref{cloudworkload} provides a description of the workloads used in our experiments. Section \ref{sec:datacenter} gives details of the data center profiling which we did to emulate real-world data centers. Section \ref{sec:sysmodel} defines the system model. Section \ref{sec:conclusion} provides details of the experiments and Section \ref{sec:eval} concludes the paper with discussion and future work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RELATED WORK}
\label{sec:related} 
Cost minimization has been extensively studied both in the context of a single data center and a geo-distributed data center. Most of the work has been targeted towards minimizing the energy consumption of the data center to reduce the operational cost. For example, the work by Gu \textit{et al.} \cite{gu2014cost} studies the influence of task assignment, data placement  and data movement on the operational expenditure of large-scale geo-distributed data centers for big data applications. The authors jointly optimize  these three factors by proposing a 2-D Markov chain to derive the average task completion time and solve the model as a MILP problem. On similar lines Agarwal \textit{et al.} \cite{agarwal2010volley} propose an automated data placement mechanism Volley for geo-distributed cloud services with the consideration of WAN bandwidth cost, data center capacity limits, data inter-dependencies, etc.

Yu \textit{et al.}\cite{yu2015energy} propose minimizing energy cost for distributed Internet data centers (IDCs) in smart microgrids by considering system dynamics like uncertainties in electricity price, workload, renewable energy generation, and power outage state. They model the problem as a stochastic program and solve it using Lyapunov optimization technique. The work by Ren and He  \cite{ren2013coca} propose an online algorithm, called COCA to minimize data center operational cost while satisfying carbon neutrality. Unlike some of the existing research, COCA enables distributed server-level resource management: each server autonomously adjusts its processing speed and optimally decides the amount of workloads to process. This work is restricted to a single data center with heterogeneous resources.

Cooling energy is a substantial part of the total energy spent by a data center. According to some reports, this can be as high as 50\% \cite{sullivan2002alternating}, \cite{patel2003smart},\cite{sawyer2004calculating} of the total energy budget. There has been a lot of work both at the application/middleware level \cite{TempLDBSC11}, \cite{leverich2010energy} as well as provisioning techniques  \cite{tang2007thermal}, \cite{chen2010integrated} that try to minimize the cooling energy.

Other research related to cooling efficiency include the work by Liu \textit{et al.} \cite{liu2012renewable}  that reduces electricity cost and environmental impact using a holistic approach that integrates renewable supply, dynamic pricing, and cooling supply including chiller and outside air cooling, with IT workload planning to improve the overall sustainability of data center operations. The authors first predict renewable energy as well as IT demand. Then they use these predictions to generate an IT workload management plan that schedules IT workload and allocates IT resources within a data center according to time varying power supply and cooling efficiency. This is done in the context of a single data center.

\iffalse
Pakbaznia and Pedram \cite{pakbaznia2009minimizing} focuses on minimizing power for both IT equipment and air conditioning power usage. The authors consider both server consolidation and task assignment together and formulate the resulting optimization problem a a ILP problem and present a heuristic algorithm that loves it in polynomial time.

There are also research done to look at the impact of failures in data centers on the operational cost. The work by Cui \textit{et al.} \cite{cui2014shadows} considers the increase in energy consumption due to failures in large scale clouds. A slower shadow replication of the main process is used to recover from faults thereby saving energy up to 30\%. Guenter \textit{et al.} \cite{guenter2011managing} also considers different trade-offs between cost, performance and reliability to formulate an optimization problem that minimizes energy costs, reliability costs and unmet demands. VM migration techniques are also used extensively to increase the energy efficiency of cloud data centers. One such approach is proposed by Rodero \textit{et al.} \cite{rodero2012energy} where the author  propose to use application profiling (CPU, memory, bandwidth) to  design application-centric energy-aware strategy for VM allocation during VM migration.

Researchers have also studied the efficient use of green energy in data centers. Goiri \textit{et al.} \cite{goiri2011greenslot} propose a system GreenSlot that maximizes the use of green energy available at data centers while ensuring all job deadlines are met. GreenSlot predicts the solar energy available in data centers in near future and schedules the workload to maximize green energy consumption and meeting jobs deadlines. Green- Slot can increase green energy consumption by up to 117\% and decrease energy cost by up to 39\%, compared to a conventional scheduler.

\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{CLOUD WORKLOADS } \label{cloudworkload}
With the increase in the adoption of cloud computing by businesses around the world, more and more workloads are moving out of traditional data centers into the cloud. A cloud data center typically handles a range of IT workloads. These workloads include service-critical and latency sensitive interactive applications that run 24x7 such as instant messengers, e-mail applications and other Internet services, and batch-style applications such as scientific HPC applications, Big Data applications like financial analysis and Map Reduce applications. A thoughtful management of these workloads are necessary not only to provide customers with low-latency services but also to reduce the energy consumption of the data centers that run these workloads. A good way to study these workloads are by analyzing real-world or synthetic workload traces. We look at cloud workload traces in detail in section \ref{traces} and \ref{googletrace}

\subsection{Workload Traces} \label{traces}
Unfortunately, there is a lack of real-world traces in the cloud computing research community. As a result, there are limited work done to understand the diversity in cloud workloads. The first large-scale workload trace was released by Google in 2011\cite{googletracedata}, featuring traces from over 12,000 servers over a period of 29 days. Various studies have been carried out to understand the characteristic of the Google trace data. Other real-world traces include Wikipedia request trace \cite{urdaneta2009wikipedia} and Hadoop logs from Yahoo! \cite{yahootrace}There have also been attempts to generate synthetic workload traces using known real-world workload characteristics \cite{beitch2010rain} \cite{wang2011towards}. For our experiments we use the workload trace from Google.

\subsection{Custom Trace Format}
To have a common trace format to represent different real-world and synthetic workload traces, we have formulated a trace format similar to standard workload format (swf) \cite{chapin1999benchmarks} defined for grid and supercomputing traces. Table \ref{tab:format} provides information about each field in the trace format. The format provides a uniform schema that can be used by different cloud providers to supply their workload traces. This can also be used as a common input format by various scheduling simulators. Our simulator, \textit{GeoSim} which is explained in section \ref{sec:simdesign} also used this format to represent jobs.

\begin {table}[]
\centering
 \scalebox{0.85}{\begin{tabular}{ | p{2.5cm} |l  |p{6 cm}|  }
    \hline
    \bf{Field Name} & \bf{Type} & \bf{Description}   \\ \hline
     JobID & INT  & Unique job identifier for each job\\ \hline
     Arrival Time & INT & Arrival time in microseconds \\ \hline
     Scheduled Time & INT  & Arrival time in microseconds\\ \hline
     Finish Time & INT  & Arrival time in microseconds\\ \hline
     Schedule class & INT  & Latency class of the job. \\ \hline
     Tasks & INT  & Total number of tasks in the job. \\ \hline
     Total CPU& INT  & Aggregate sum of CPU used by all the tasks in the job\\ \hline
     Total Memory& INT   & Aggregate sum of memory used by all the tasks in the job \\ \hline
     User& INT  & Unique user identifier \\ \hline
     SLA Param1&  FLOAT & SLA parameter placeholder\\ \hline
     SLA Param2&  FLOAT & SLA parameter placeholder\\ \hline
     SLA Param3&  FLOAT & SLA parameter placeholder\\ \hline
    \end{tabular}}
    \caption {Trace format description} \label{tab:format} 
\end{table}

\subsection{Google Workload Trace} \label{googletrace}
Google trace has job and task information in different files. We consolidated all the required information and generated a single work trace file with the format described in the previous section. For our work, we model five different data centers located at different parts of the globe. A detailed explanation about the data center profile are given in section \ref{sec:datacenter}. To simulate workload execution in these five data centers we have extracted five set of traces each worth five days of incoming job requests from the original Google trace. This will cover 25 days of traces out of the 29 available days. We label these 5 set of traces from A to E.   Additionally,  in our  experiments  we  only  consider  those  jobs  that end  without  any  failures.  Therefore,  we  clean  the original  traces  by  removing  all  jobs  that  have  either failed or got killed. The left over 4 days of the workload is used to gain statistical user related insights to predict the running time of incoming jobs.

\begin{figure}[] 
\centering
\includegraphics[scale=0.3]{arrivalrate.png}
\caption{Arrival rate of jobs }
\label{fig:arrivalrate}
\end{figure}

Now we will look at some characteristics of the traces. Table \ref{tab:tracestat} describes the statistics for each of the five traces grouped by the scheduling class of the job. In Google traces, CPU and memory usage are normalized relative to the largest capacity of the resource on any machine in the trace (which is 1.0). Most of the class 0 and class 1 jobs are batch jobs with less latency requirements. Generally, these are the kind of jobs that can be provisioned on other data centers as their deadlines are not strict.  You will find that as the sensitivity increases the running time of these jobs and the resource usage also increase. High-sensitive jobs are generally service jobs and run of longer periods of time. Figure \ref{fig:pie} shows a pie chart of the job counts in different schedule classes for all the 5 traces. You will notice that the job count for class 3 (high latency-sensitive) jobs are low. This is because these jobs generally run for long period of time, typically months. Since we only consider jobs that start and end within the 29 day range such jobs are missed out in our traces. For our work, we will consider class-0 and class-1 jobs as candidates for optimizing the provisioning. Class-2 and class-3 jobs are provisioned to run on the data center where the job arrived since these are high-sensitive and their SLA demands will be strict.

\begin{figure*}[]
\centering
\includegraphics[scale=0.65]{pie}
\caption{ Distribution of job count in various schedule classes}
 \label{fig:pie}
\end{figure*}

\begin{table*}[]
\adjustbox{max width=1\textwidth}{

\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
\hline
 & \multicolumn{3}{l|}{\bf{Trace A}} & \multicolumn{3}{l|}{\bf{Trace B}} & \multicolumn{3}{l|}{\bf{Trace C}} & \multicolumn{3}{l|}{\bf{Trace D}} & \multicolumn{3}{l|}{\bf{Trace E}} \\ \hline
\textbf{\begin{tabular}[c]{@{}l@{}}Sched\\ Class\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg \\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} \\ \hline
0 & 0.075 & 0.051 & 415 & 0.123 & 0.096 & 345 & 0.083 & 0.065 & 385 & 0.093 & 0.092 & 373 & 0.093 & 0.074 & 333 \\ \hline
1 & 0.087 & 0.029 & 729 & 0.086 & 0.032 & 760 & 0.082 & 0.028 & 711 & 0.073 & 0.030 & 743 & 0.093 & 0.043 & 693 \\ \hline
2 & 0.067 & 0.043 & 1028 & 0.096 & 0.058 & 872 & 0.088 & 0.057 & 918 & 0.073 & 0.045 & 707 & 0.084 & 0.049 & 832 \\ \hline
3 & 0.088 & 0.12 & 1105 & 0.099 & 0.12 & 82 & 0.101 & 0.012 & 7961 & 0.081 & 0.051 & 17135 & 0.089 & 0.061 & 9021 \\ \hline
\end{tabular}
}
\caption {Trace set statistics} \label{tab:tracestat} 
\end{table*}

Figure \ref{fig:arrivalrate} shows the variation in job arrival rates for the traces. As you can see there is both intra-trace variation as well as inter-trace variation. This is representative of the real-world cloud workload pattern where bursty arrivals are common. There are opportunities to provision jobs to other data centers if the traffic at a particular data center is high. What is also important to notice is that the scheduling decision should be made quickly when the arrival rates of jobs are high, else there are chances for that these jobs miss the SLAs. Hence, even the scheduler should be scalable during high volume traffic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{DATA CENTER PROFILING } \label{sec:datacenter}
In order to run our experiments, we need to collect information about data centers which will help us make better scheduling decisions. We call this process of collecting the required information about data centers as \textit{data center profiling}. The major energy costs in data center are from running the IT resources and from cooling the data centers. We create a profile for each of the data center with the information that effects the energy costs.

\subsection{Geo-Distributed Data Center}
Most cloud service providers including Google, Amazon, Facebook and Apple have data centers distributed across the globe. For our experiments, we consider five locations where Google has its data centers\cite{googlelocation}. Table \ref{tab:dc} provides the location and labels for each of the data center. We also replicate the computing resource used in the original Google trace across all the data centers. The configuration of the resources along with the normalized CPU and memory are given in table \ref{tab:dcres}

\subsection{Profile Parameters}
We consider two major parameters that has the most effect on the energy cost. They are a) Temperature b) Electricity. We collect temperature and electricity prices for these locations from historical archives. We look at each of these parameters and how they effect energy costs in sections \ref{sec:temp} and \ref{sec:elec}.

\begin{table}[] 
 \centering
\begin{tabular}{|p{3cm}|l|p{2.5cm}|}
\hline
\bf{Location}                        & \bf{Label} & \bf{Associated Trace} \\ \hline
Council Bluffs, Iowa & Iowa   &  A               \\ \hline
The Dalles, Oregon          & Oregon   & B                \\ \hline
Singapore                       & Singapore   & C                \\ \hline
Quilicura, Chile                & Chile   & D                \\ \hline
Hamina, Finland                 & Finland   &   E               \\ \hline
\end{tabular}
\caption {Data center location information} \label{tab:dc}
\end{table}

\begin{table}[] 
 \centering
\begin{tabular}{|l|l|l|l|}
\hline
\bf{Node count}      &       \bf{Architecture}          & \bf{CPU} & \bf{Memory} \\ \hline
6732 & B & 0.5 & 0.50   \\ \hline
3863 & B & 0.5 & 0.25   \\ \hline
1001 & B & 0.5 & 0.75   \\ \hline
795 & C & 1 & 1   \\ \hline
126 & A & 0.25 & 0.25   \\ \hline
52 & B & 0.5 & 0.12  \\ \hline
5 & B & 0.5 & 0.3  \\ \hline
5 & B & 0.5 & 0.97  \\ \hline
3 & C & 1 & 0.50  \\ \hline
1 & B & 0.5 & 0.06  \\ \hline
\end{tabular}
\caption {Data center IT resource configuration} \label{tab:dcres}
\end{table}

\begin{figure}[H]
\includegraphics[scale=0.45]{temperature}
\caption{ Temperature variation for the year 2013}
 \label{fig:temp}
\end{figure}

\subsubsection{Temperature} \label{sec:temp}
Data centers located in cold regions have the advantage of using outside air to cool the data center. An air-side economizer brings outside air into a building and distributes it to the servers. Instead of being re-circulated and cooled, the exhaust air from the servers is simply directed outside. This can result in million dollars of saving annually. A difference in temperature across data centers leads to different cooling costs at these data centers. We wish to utilize this difference by scheduling jobs in low temperature data centers. In this case we make our assumption concrete by doing an empirical analysis of historical climate data. We extracted hourly temperature data from various data repositories of the National Climate Data Center \cite{climatedata} for all five locations covering the entire one-year period of 2013. Figure \ref{fig:temp} plots the weekly average temperature for five different locations in United States, Europe, South-America, and Asia. As expected we see high diversity in temperature at different locations. For example, Oregon and Finland seems to be equally favorable for cooling during most of the year except during the months of May {-} July. Chile and Singapore show almost no variation in temperature across the year. However, Chile is favorable for cooling conditions compared to Singapore.

\subsubsection{Electricity Pricing} \label{sec:elec}
Creation of power trading zones have resulted in wholesale power purchase under dynamic pricing schemes. Such markets are popular in USA \cite{elecusa}, Germany \cite{elecgermany} and other countries. Based on supply and demand bids there can be high fluctuations in prices \cite{benini2002day}. There are many ways to predict these prices \cite{contreras2003arima} \cite{garcia2005garch}. If we can predict the electricity prices for the job duration, then we will be able to schedule the job such that the running costs go down. Figure \ref{fig:elec} shows the variation in electricity prices across different data center regions in 2013. These price values have been consolidated from multiple sources on the Internet \cite{singelec} \cite{finelec} \cite{uselec}. For Chile, we were able to retrieve only monthly averages for 2013, but for all the other location we were successful in collecting hourly prices. From the figure \ref{fig:elec} it is clear that Singapore has the highest electricity pricing among all the regions under consideration. If we compare both figure \ref{fig:temp} and figure \ref{fig:elec} it is evident that Singapore is least suitable data center in terms of operational costs since the temperature and the energy prices are the highest.

\begin{figure}[H]
\includegraphics[scale=0.45]{elec}
\caption{ Electricity price variation for the year 2013}
 \label{fig:elec}
\end{figure}

\section{System Model}
\label{sec:sysmodel}
We model the total energy of the data center as a sum of energy consumed by the servers i.e IT equipments and the cooling subsystem. The subsection \ref{sec:servere} and \ref{sec:coole} provides more information about the model.

\begin{equation} \label{eq:etotal}
E_{total} = E_{server} + E_{cooling} 
\end{equation}

\subsection{Server Energy}
\label{sec:servere}
We model the power consumption of servers as the sum of power consumption of individual cores in the system. The power consumption on each core is the sum of the static power and the dynamic power needed to run computations. Eq \ref{eq:core} represents the power consumption of a core. Static power is consumed when the system is idle.

\begin{equation} \label{eq:core}
P_{core} = P_{static} + P_{dynamic} 
\end{equation}

The peak usage is when the cores are at the highest utilization. In cloud systems, it is very rare that the CPU utilization \textit{u} is more than 20\% \cite{birke2012data}. To account for this, we rewrite eq \ref{eq:core} as described in \cite{buyya2010energy}.

\begin{equation} \label{eq:core_util}
P_{core}(u) = P_{static} + P_{dynamic_peak} * u 
\end{equation}

For our experiments we use the peak and idle power from \cite{kamil2008power}. The authors have characterized power usage of systems by running various HPC kernels. Since these kernels are CPU-intensive we can consider them to use the peak power. To emulate the CPU utilization of jobs, we generate random utilization based on the \textit{mean, standard deviation} and \textit{cov} provided in the work done on data center utilization by IBM \cite{birke2012data}.

The energy consumption of a job \textit{j} with core count \textit{c}, avgerage CPU utilization \textit{u} and running time \textit{t} is given in Eq \ref{eq:energy_job}

\begin{equation} \label{eq:power_job}
P_{j} = c*P_{core}(u)
\end{equation}

\begin{equation} \label{eq:energy_job}
E_{j} = P_{j}*t
\end{equation}

\subsection{Cooling Energy}
\label{sec:coole}
We model the energy consumption of the traditional CRAC as well as the air economizer in the sections below.

\subsubsection{CRAC cooling units}
The power usage of CRAC units can be modeled using the equation  \ref{crac:eq} as given in \cite{christy2011energy}. The value of COP (coefficient of performance) can be calculated according to \cite{ahmad2010joint} using the equation \ref{eq:cop}.

\begin{equation} \label{crac:eq}
P_{CRAC} = \frac{P_{dc}}{COP}
\end{equation}

\begin{equation} \label{eq:cop}
COP = 0.0068*(T_{sup}^2) + 0.0008*T_{sup} + 0.458
\end{equation}

In the above equations, ~$P_{dc}$ is the total power consumption of the data center and ~$T_{sup}$ is the temperature of the air supplied to the data center in degree Celsius.

\subsubsection{Air Economizer}

The power usage model for air economizer is given in \cite{christy2011energy}. The equation \ref{eq:aireco} represents the power consumption of the free air cooler.

\begin{equation} \label{eq:aireco}
P_{air} = (PUE - 1) * P_{dc}
\end{equation}

The PUE(power usage effectiveness) is proportional to the outside temperature. We get the values of PUE from the product documentation of commercial vendors \cite{airecoproduct}. For our experiments, the value of PUE used is given in table \ref{tab:pue}.

\begin{table}[] 
 \centering
\begin{tabular}{|l|l|}
\hline
\bf{Temperature range (F)}      &       \bf{PUE} \\ \hline
less than 25  & 1.05    \\ \hline
25-35 & 1.07   \\ \hline
35-50 & 1.09    \\ \hline
50-60 & 1.10  \\ \hline
60-65 & 1.17  \\ \hline
\end{tabular}
\caption {PUE values for different temperature ranges} \label{tab:pue}
\end{table}

\subsection{Data center selection}
The data center is selected by calculating the power contributed by a job to the total server power consumption as well as cooling power. This is done by using Eq \ref{eq:power_job} in the power models for server and cooling. 
Equations \ref{eq:total_job_power}, \ref{eq:job_cost} and \ref{eq:job_min} show the mechanism of data center selection.

\begin{equation} \label{eq:total_job_power}
P_{total}(j,dc) = P_{dynamic}(j,dc) + P_{cooling}(j,dc)
\end{equation}

\begin{equation} \label{eq:job_cost}
Cost(j,dc) = T_{j}*P_{total}(j,dc)*C_{mwh}(dc)
\end{equation}

\begin{equation} \label{eq:job_min}
dc_{j} = \min_{\forall dc \in DC} Cost(j,dc)
\end{equation}

The cost is the product of total power due to job ~$j$ in a specific data center ~$dc$, the running time of the job ~$T(j)$ and the electricity cost at the data center ~$C_{mwh}(dc)$. The data center corresponding to the minimum cost is selected for job scheduling as specified in Eq \ref{eq:job_min}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Evaluation}
\label{sec:eval}
\subsection{Simulator}
\label{sec:simdesign}
We evaluate our meta-scheduler by doing extensive simulations. For this purpose we developed a simulator \textit{GeoSim} which take in the workload trace and the data center profile as input and implements our model and routes jobs to the most cost effective data center. A brief overview of the design of \textit{GeoSim} is provided below. We have made the source code of \textit{GeoSim} and the workload traces public and is available at \cite{sourcecode}

\

\subsubsection{Components}
\begin{list2} %Job Description%
\item \textbf{Datacenter:} This component will simulate a data center. This composed of 1) scheduler that implements a specific provisioning algorithm, 2) data center profile database, 3) a work trace reader and 4)proxy elements of other data center for inter-data center communication\
\item \textbf{DataCenter Proxy}: All interactions with the data center will be through a proxy. Each data center will hold the proxies to other data centers.\
\item \textbf{Scheduler:} This component will implement the model and schedule jobs onto data centers. We can experiment with different scheduling algorithms by plugging different Scheduler implementation\
\item \textbf{Datacenter Profile DB}: This components contains the temperature and electricity data. Scheduler will access this to make provisioning decisions.\
\end{list2}

\begin{figure}[] 
\centering
\includegraphics[scale=0.45]{comp_diag}
\caption{Simulation components and their interaction}
\label{fig:component}
\end{figure}

\subsubsection{Component Interaction}
Figure \ref{fig:component} gives the interaction model between the components. The data center object will start a thread which will run the scheduling algorithm, scheduling jobs from the trace file. Each data center has proxies to other data centers if it needs to communicate resource availability or job provisioning to other data centers. In our simulations we share resource utilization every 5 minutes, and each time step of the simulator is 5 seconds. These are provided as parameters to the simulator. Our simulations take around 4 hours to complete running on 6-core AMD Opteron(tm) server.

\begin{figure}[H]
\includegraphics[scale=0.45]{cpu_util}
\caption{ Data center utilization}
 \label{fig:dcutil}
\end{figure}

\subsection{Results}
We run the simulations with GeoSched and also implement a naive scheduler called \textit{DumbSched} which schedules all the jobs to the local data centers. We will look at the cost saving as well as the energy savings with GeoSched. We will also look at the variation in utilization of each of the data centers across different months. We run the simulations for alternate months starting with January. We provide the date 15 and GMT time 12am as the start date and time for each simulation.

The results are summarized in table \ref{table:summary}. The total energy consumed by the data centers for 30 days (5 days in each of the 6 months) and the corresponding costs are shown. The energy  decrease by 8.2\% and the cost decrease by 11.7\%. When these results are extrapolated to 1 year the dollars  saved is around \$4.8 million. The experiments were run assuming a max core count in node to be 64. Since, Google traces normalize these values, it would be impossible to calculate energy with out an assumption.

\begin{table}[]

 \centering
\begin{tabular}{|l|l|l|}
\hline
      &       \bf{Energy (MJ)} & \bf{Cost (\$)} \\ \hline
\bf{DumbSched} & 8044  & 560630   \\ \hline
\bf{GeoSched} & 7384 & 494560   \\ \hline
\bf{ Savings! } & 660 & 66070    \\ \hline
\end{tabular}
\caption {Result summary} \label{table:summary}
\end{table}

Figure \ref{fig:dcutil} shows the variation of data center utilization when the scheduling month changes. It is quiet evident that Singapore has the least utilization due to high operational costs. Similarly, the utilization at Iowa decreases towards may and increase towards the end of the year. This result is consistent with the temperature and electricity price data. Finland and Oregon has the highest utilization since these two have less operational cost as compared to the other three data centers. Chile, gets more jobs towards the middle of the year. This could be attributed to temperature increase in other data centers.

Figure \ref{fig:dcenergy} shows the variation in energy consumption. It can be noted that some data centers consume more energy with GeoSched than DumbSched. This is because, in these data centers, more work can be done with less increase in energy consumption. Therefore, GeoSched will give a net decrease in energy consumption even if some data centers consume more energy than when scheduled using DumbSched.

\begin{figure}[H]
\includegraphics[scale=0.45]{energy}
\caption{ Variation in energy consumption}
 \label{fig:dcenergy}
\end{figure}

Figure \ref{fig:dccost} shows the cost of operation among different data centers. This plot can also be interpreted the same way as the energy consumption plot. There is a net gain in the cost of operating the data center when using GeoSched. More energy can be saved if the free nodes in the data center can be switched off. In our experiments, we always calculate the idle energy consumed by these machines.

\begin{figure}[]
\includegraphics[scale=0.45]{dollars}
\caption{ Variation in operational cost}
 \label{fig:dccost}
\end{figure}

\section{CONCLUSION}
\label{sec:conclusion}
In this paper we build a meta-scheduler for geo-distributed data centers. We considered parameters like temperature and electricity prices to intelligently provision jobs to data centers. We evaluate our method using simulations and found our scheduler to reduce energy consumed by 8.2\% and the cost of operation 11.7\%. We wish to consider other factor influencing the energy consumption like availability of green energy and data transfer energy for our future work. We would also like to consider more descriptive SLAs to be included as part of our future work.

{\footnotesize \bibliographystyle{acm}
\bibliography{sample}}

% that's all folks
\end{document}
